====== Function PEEK-CHAR ======

====Syntax====
  * **peek-char** ''&optional'' //peek-type// //input-stream// //eof-error-p// //eof-value// //recursive-p// → //char//

====Arguments and Values====
  * //peek-type// - a //[[CL:Glossary:character]]// or **[[CL:Constant Variables:t]]** or **[[CL:Constant Variables:nil]]**.
  * //input-stream// - //[[CL:Glossary:input]]// //[[CL:Glossary:stream designator]]//. The default is //[[CL:Glossary:standard input]]//.
  * //eof-error-p// - a //[[CL:Glossary:generalized boolean]]//. The default is //[[CL:Glossary:true]]//.
  * //eof-value// - an //[[CL:Glossary:object]]//. The default is **[[CL:Constant Variables:nil]]**.
  * //recursive-p// - a //[[CL:Glossary:generalized boolean]]//. The default is //[[CL:Glossary:false]]//.
  * //char// - a //[[CL:Glossary:character]]// or the //eof-value//.

====Description====
**peek-char** obtains the next character in //input-stream// without actually reading it, thus leaving the character to be read at a later time. It can also be used to skip over and discard intervening characters in the //input-stream// until a particular character is found.

If //peek-type// is not supplied or **[[CL:Constant Variables:nil]]**, **peek-char** returns the next character to be read from //input-stream//, without actually removing it from //input-stream//. The next time input is done from //input-stream//, the character will still be there.

If //peek-type// is **[[CL:Constant Variables:t]]**, then **peek-char** skips over //[[CL:Glossary:whitespace]]// //[[CL:Glossary:characters]]//, but not comments, and then performs the peeking operation on the next character. The last character examined, the one that starts an //[[CL:Glossary:object]]//, is not removed from //input-stream//.

If //peek-type// is a //[[CL:Glossary:character]]//, then **peek-char** skips over input characters until a character that is **[[CL:Functions:char=]]** to that //[[CL:Glossary:character]]// is found; that character is left in //input-stream//.

If an //[[CL:Glossary:end of file]]// occurs and //eof-error-p// is //[[CL:Glossary:false]]//, //eof-value// is returned.

If //recursive-p// is //[[CL:Glossary:true]]//, this call is expected to be embedded in a higher-level call to **[[CL:Functions:read]]** or a similar //[[CL:Glossary:function]]// used by the //[[CL:Glossary:Lisp reader]]//.

When //input-stream// is an //[[CL:Glossary:echo stream]]//, characters that are only peeked at are not echoed. In the case that //peek-type// is not **[[CL:Constant Variables:nil]]**, the characters that are passed by **peek-char** are treated as if by **[[CL:Functions:read-char]]**, and so are echoed unless they have been marked otherwise by **[[CL:Functions:unread-char]]**.

====Examples==== 
<blockquote> 
([[CL:Macros:with-input-from-string]] (input-stream " 1 2 3 4 5") 
  ([[CL:Functions:format]] [[CL:Constant Variables:t]] "~S ~S ~S" 
          (peek-char t input-stream) 
          (peek-char #\4 input-stream) 
          (peek-char nil input-stream)))
<o>#\1 #\4 #\4 </o>
<r>NIL </r>
</blockquote>

====Affected By====
**[[CL:Variables:star-readtable-star|*readtable*]]**, **[[CL:Variables:star-standard-input-star|*standard-input*]]**, 
**[[CL:Variables:star-terminal-io-star|*terminal-io*]]**.

====Exceptional Situations====
If //eof-error-p// is //[[CL:Glossary:true]]// and an //[[CL:Glossary:end of file]]// occurs an error of type **[[CL:Types:end-of-file]]** is signaled.

If //peek-type// is a //[[CL:Glossary:character]]//, an //[[CL:Glossary:end of file]]// occurs, and //eof-error-p// is //[[CL:Glossary:true]]//, an error of type **[[CL:Types:end-of-file]]** is signaled.

If //recursive-p// is //[[CL:Glossary:true]]// and an //[[CL:Glossary:end of file]]// occurs, an error of type **[[CL:Types:end-of-file]]** is signaled.

====See Also====
None.

====Notes====
None.

\issue{ARGUMENTS-UNDERSPECIFIED:SPECIFY} \issue{PEEK-CHAR-READ-CHAR-ECHO:FIRST-READ-CHAR}
