====== Special Operator TAGBODY ======

====Syntax====

\DefspecWithValues tagbody {\star{{//tag// | //statement//}}} {**[[CL:Constant Variables:nil]]**}

====Arguments and Values====
  * //tag// - a //[[CL:Glossary:go tag]]//; not evaluated.
  * //statement// - a //[[CL:Glossary:compound form]]//; evaluatedspecial.

====Description====

Executes zero or more //[[CL:Glossary:statements]]// in a

//[[CL:Glossary:lexical environment]]// that provides for control transfers to labels indicated by the //tags//.

The //statements// in a **[[CL:Special Operators:tagbody]]** are //[[CL:Glossary:evaluate|evaluated]]// in order from left to right, and their //[[CL:Glossary:value|values]]// are discarded. If at any time there are no remaining //statements//, **[[CL:Special Operators:tagbody]]** returns **[[CL:Constant Variables:nil]]**.

However, if ''(go //tag//)'' is //[[CL:Glossary:evaluate|evaluated]]//, control jumps to the part of the body labeled with the //tag//. (Tags are compared with **[[CL:Functions:eql]]**.)

A //tag// established by **[[CL:Special Operators:tagbody]]** has //[[CL:Glossary:lexical scope]]// and has //[[CL:Glossary:dynamic extent]]//. Once **[[CL:Special Operators:tagbody]]** has been exited, it is no longer valid to **[[CL:Special Operators:go]]** to a //tag// in its body. It is permissible for **[[CL:Special Operators:go]]** to jump to a **[[CL:Special Operators:tagbody]]** that is not the innermost **[[CL:Special Operators:tagbody]]** containing that **[[CL:Special Operators:go]]**; the //tags// established by a **[[CL:Special Operators:tagbody]]** only shadow other //tags// of like name.

The determination of which elements of the body are //tags// and which are //statements// is made prior to any //[[CL:Glossary:macro expansion]]// of that element. If a //statement// is a //[[CL:Glossary:macro form]]// and its //[[CL:Glossary:macro expansion]]// is an //[[CL:Glossary:atom]]//, that //[[CL:Glossary:atom]]// is treated as a //statement//, not a //tag//.

====Examples====

<blockquote> (let (val) (tagbody ([[CL:Macros:defparameter]] val 1) (go point-a) (incf val 16) point-c (incf val 04) (go point-b) (incf val 32) point-a (incf val 02) (go point-c) (incf val 64) point-b (incf val 08)) val) → 15 (defun f1 (flag) (let ((n 1)) (tagbody ([[CL:Macros:defparameter]] n (f2 flag #'(lambda () (go out)))) out (prin1 n)))) → F1 (defun f2 (flag escape) (if flag (funcall escape) 2)) → F2 (f1 nil)
▷ 2 → NIL (f1 t)
▷ 1 → NIL </blockquote>

====Affected By====

None.

====Exceptional Situations====

None.

====See Also====

**[[CL:Special Operators:go]]**

====Notes====

The //[[CL:Glossary:macro|macros]]// in \thenextfigure\ have //[[CL:Glossary:implicit tagbodies]]//.

\displaythree{Macros that have implicit tagbodies.}{ do&do-external-symbols&dotimes\cr do*&do-symbols&prog\cr do-all-symbols&dolist&prog*\cr }

\issue{TAGBODY-TAG-EXPANSION:NO}
