====== Special Operator LET, LET* ======

====Syntax====

\DefspecWithValues let {\paren{\star{\VarValue}} \starparam{declaration} \starparam{form}} {\starparam{result}}

\DefspecWithValues let* {\paren{\star{\VarValue}} \starparam{declaration} \starparam{form}} {\starparam{result}}

====Arguments and Values====

//var// - a //[[CL:Glossary:symbol]]//.

//init-form// - a //[[CL:Glossary:form]]//.

//declaration// - a **[[CL:Special Operators:declare]]** //[[CL:Glossary:expression]]//; not evaluated.

//form// - a //[[CL:Glossary:form]]//.

//results// - the //[[CL:Glossary:value|values]]// returned by the //[[CL:Glossary:forms]]//.

====Description====

**[[CL:Special Operators:let]]** and **[[CL:Special Operators:let*]]** create new variable //[[CL:Glossary:binding|bindings]]// and execute a series of //forms// that use these //[[CL:Glossary:binding|bindings]]//. **[[CL:Special Operators:let]]** performs the //[[CL:Glossary:binding|bindings]]// in parallel and **[[CL:Special Operators:let*]]** does them sequentially.


The form

<blockquote> (let ((//var1// //init-form-1//) (//var2// //init-form-2//) ... (//varm// //init-form-m//)) //declaration1// //declaration2// ... //declarationp// //form1// //form2// ... //formn//) </blockquote> first evaluates the expressions //init-form-1//, //init-form-2//, and so on,

in that order, saving the resulting values. Then all of the variables //varj// are bound to the corresponding values; each //[[CL:Glossary:binding]]// is lexical unless there is a **[[CL:Declarations:special]]** declaration to the contrary. The expressions //formk// are then evaluated in order; the values of all but the last are discarded (that is, the body of a **[[CL:Special Operators:let]]** is an //[[CL:Glossary:implicit progn]]//).

**[[CL:Special Operators:let*]]** is similar to **[[CL:Special Operators:let]]**, but the //[[CL:Glossary:binding|bindings]]// of variables are performed sequentially rather than in parallel. The expression for the //init-form// of a //var// can refer to //vars// previously bound in the **[[CL:Special Operators:let*]]**.


The form

<blockquote> (let* ((//var1// //init-form-1//) (//var2// //init-form-2//) ... (//varm// //init-form-m//)) //declaration1// //declaration2// ... //declarationp// //form1// //form2// ... //formn//) </blockquote> first evaluates the expression //init-form-1//, then binds the variable //var1// to that value; then it evaluates //init-form-2// and binds

//var2//, and so on. The expressions //formj// are then evaluated in order; the values of all but the last are discarded (that is, the body of **[[CL:Special Operators:let*]]** is an implicit **[[CL:Special Operators:progn]]**).

For both **[[CL:Special Operators:let]]** and **[[CL:Special Operators:let*]]**, if there is not an //init-form// associated with a //var//, //var// is initialized to **[[CL:Constant Variables:nil]]**.

The special form **[[CL:Special Operators:let]]** has the property that the //[[CL:Glossary:scope]]// of the name binding does not include any initial value form. For **[[CL:Special Operators:let*]]**, a variable's //[[CL:Glossary:scope]]// also includes the remaining initial value forms for subsequent variable bindings.

====Examples====

<blockquote> ([[CL:Macros:defparameter]] a 'top) → TOP (defun dummy-function () a) → DUMMY-FUNCTION (let ((a 'inside) (b a)) (format nil "~S ~S ~S" a b (dummy-function))) → "INSIDE TOP TOP" (let* ((a 'inside) (b a)) (format nil "~S ~S ~S" a b (dummy-function))) → "INSIDE INSIDE TOP" (let ((a 'inside) (b a)) (declare (special a)) (format nil "~S ~S ~S" a b (dummy-function))) → "INSIDE TOP INSIDE" </blockquote>

\medbreak The code

<blockquote> (let (x) (declare (integer x)) ([[CL:Macros:defparameter]] x (gcd y z)) ...) </blockquote> is incorrect; although ''x'' is indeed set before it is used, and is set to a value of the declared type //[[CL:Glossary:integer]]//, nevertheless ''x'' initially takes on the value **[[CL:Constant Variables:nil]]** in violation of the type declaration.

====Affected By====

None.

====Exceptional Situations====

None.

====See Also====

**[[CL:Special Operators:progv]]**

====Notes====

None.

\issue{DECLS-AND-DOC} \issue{VARIABLE-LIST-ASYMMETRY:SYMMETRIZE} \issue{KMP-COMMENTS-ON-SANDRA-COMMENTS:X3J13-MAR-92} \issue{KMP-COMMENTS-ON-SANDRA-COMMENTS:X3J13-MAR-92}
