====== Variable *PRINT-LEVEL*, *PRINT-LENGTH* ======

====Value Type====
a non-negative //[[CL:Glossary:integer]]//, or **[[CL:Constant Variables:nil]]**.

====Initial Value====
**[[CL:Constant Variables:nil]]**.

====Description====
**<nowiki>*print-level*</nowiki>** controls how many levels deep a nested //[[CL:Glossary:object]]// will print. If it is //[[CL:Glossary:false]]//, then no control is exercised. Otherwise, it is an //[[CL:Glossary:integer]]// indicating the maximum level to be printed. An //[[CL:Glossary:object]]// to be printed is at level ''0''; its components (as of a //[[CL:Glossary:list]]// or //[[CL:Glossary:vector]]//) are at level ''1''; and so on. If an //[[CL:Glossary:object]]// to be recursively printed has components and is at a level equal to or greater than the //[[CL:Glossary:value]]// of **<nowiki>*print-level*</nowiki>**, then the //[[CL:Glossary:object]]// is printed as ""#''''.

**<nowiki>*print-length*</nowiki>** controls how many elements at a given level are printed. If it is //[[CL:Glossary:false]]//, there is no limit to the number of components printed. Otherwise, it is an //[[CL:Glossary:integer]]// indicating the maximum number of //[[CL:Glossary:element|elements]]// of an //[[CL:Glossary:object]]// to be printed. If exceeded, the printer will print ''...'' in place of the other //[[CL:Glossary:element|elements]]//. In the case of a //[[CL:Glossary:dotted list]]//, if the //[[CL:Glossary:list]]// contains exactly as many //[[CL:Glossary:element|elements]]// as the //[[CL:Glossary:value]]// of **<nowiki>*print-length*</nowiki>**, the terminating //[[CL:Glossary:atom]]// is printed rather than printing ''...''

**<nowiki>*print-level*</nowiki>** and **<nowiki>*print-length*</nowiki>** affect the printing of an any //[[CL:Glossary:object]]// printed with a list-like syntax. They do not affect the printing of //[[CL:Glossary:symbols]]//, //[[CL:Glossary:strings]]//, and //[[CL:Glossary:bit vectors]]//.

====Examples====
<blockquote> 
([[CL:Special Operators:let]] ((a '(1 (2 (3 (4 (5 (6))))))))
  ([[CL:Macros:dotimes]] (i 8) 
    ([[CL:Special Operators:let]] ((*print-level* i)) 
      ([[CL:Functions:format]] [[CL:Constant Variables:t]] "~&~D -- ~S~%" i a))))
<o>0 -- #
1 -- (1 #)
2 -- (1 (2 #))
3 -- (1 (2 (3 #)))
4 -- (1 (2 (3 (4 #))))
5 -- (1 (2 (3 (4 (5 #)))))
6 -- (1 (2 (3 (4 (5 (6))))))
7 -- (1 (2 (3 (4 (5 (6)))))) </o>
<r>NIL </r>
([[CL:Special Operators:let]] ((a '(1 2 3 4 5 6)))
  ([[CL:Macros:dotimes]] (i 7) 
    ([[CL:Special Operators:let]] ((*print-length* i)) 
      ([[CL:Functions:format]] [[CL:Constant Variables:t]] "~&~D -- ~S~%" i a))))
<o>0 -- (...)
1 -- (1 ...)
2 -- (1 2 ...)
3 -- (1 2 3 ...)
4 -- (1 2 3 4 ...)
5 -- (1 2 3 4 5 6)
6 -- (1 2 3 4 5 6) </o>
<r>NIL </r>
([[CL:Macros:dolist]] (level-length '((0 1) (1 1) (1 2) (1 3) (1 4) (2 1) (2 2) (2 3) (3 2) (3 3) (3 4))) 
  ([[CL:Special Operators:let]] ((*print-level* ([[CL:Functions:first]] level-length)) 
        (*print-length* ([[CL:Functions:second]] level-length)))
    ([[CL:Functions:format]] [[CL:Constant Variables:t]] "~&~D ~D -- ~S~%"
            *print-level* *print-length* 
            '([[CL:Special Operators:if]] ([[CL:Functions:member]] x y) ([[CL:Functions:math-add|+]] ([[CL:Functions:car]] x) 3) '(foo . #(a b c d "Baz"))))))
<o>0 1 -- #
1 1 -- ([[CL:Special Operators:IF]] ...)
1 2 -- ([[CL:Special Operators:IF]] # ...)
1 3 -- ([[CL:Special Operators:IF]] # # ...)
1 4 -- ([[CL:Special Operators:IF]] # # #)
2 1 -- ([[CL:Special Operators:IF]] ...)
2 2 -- ([[CL:Special Operators:IF]] ([[CL:Functions:MEMBER]] X ...) ...)
2 3 -- ([[CL:Special Operators:IF]] ([[CL:Functions:MEMBER]] X Y) ([[CL:Functions:math-add|+]] # 3) ...)
3 2 -- ([[CL:Special Operators:IF]] ([[CL:Functions:MEMBER]] X ...) ...)
3 3 -- ([[CL:Special Operators:IF]] ([[CL:Functions:MEMBER]] X Y) ([[CL:Functions:math-add|+]] ([[CL:Functions:CAR]] X) 3) ...)
3 4 -- ([[CL:Special Operators:IF]] ([[CL:Functions:MEMBER]] X Y) ([[CL:Functions:math-add|+]] ([[CL:Functions:CAR]] X) 3) '(FOO . #(A B C D ...))) </o>
<r>NIL </r>
</blockquote>

====Affected By====
None.

====See Also====
  * **[[CL:Functions:write|Function WRITE]]**

====Notes====
None.

