====== Macro DEFCONSTANT ======

====Syntax====
  * **defconstant** //name initial-value [documentation]// → //name//

====Arguments and Values====
  * //name// - a //[[CL:Glossary:symbol]]//; not evaluated.
  * //initial-value// - a //[[CL:Glossary:form]]//; evaluated.
  * //documentation// - a //[[CL:Glossary:string]]//; not evaluated.

====Description====
**defconstant** causes the global variable named by //name// to be given a value that is the result of evaluating //initial-value//.

A constant defined by **defconstant** can be redefined with **defconstant**. However, the consequences are undefined if an attempt is made to assign a //[[CL:Glossary:value]]// to the //[[CL:Glossary:symbol]]// using another operator, or to assign it to a //[[CL:Glossary:different]]// //[[CL:Glossary:value]]// using a subsequent **defconstant**.

If //documentation// is supplied, it is attached to //name// as a //[[CL:Glossary:documentation string]]// of kind **[[CL:Types:variable]]**.

**defconstant** normally appears as a //[[CL:Glossary:top level form]]//, but it is meaningful for it to appear as a //[[CL:Glossary:non-top-level form]]//. However, the compile-time side effects described below only take place when **defconstant** appears as a //[[CL:Glossary:top level form]]//.

The consequences are undefined if there are any //[[CL:Glossary:binding|bindings]]// of the variable named by //name// at the time **defconstant** is executed or if the value is not **[[CL:Functions:eql]]** to the value of //initial-value//.

The consequences are undefined when constant //[[CL:Glossary:symbols]]// are rebound as either lexical or dynamic variables. In other words, a reference to a //[[CL:Glossary:symbol]]// declared with **defconstant** always refers to its global value.

The side effects of the execution of **defconstant** must be equivalent to at least the side effects of the execution of the following code:

<blockquote>
([[CL:Macros:setf]] (symbol-value '//name//) //initial-value//)
([[CL:Macros:setf]] (documentation '//name// 'variable) '//documentation//)
</blockquote>

If a **defconstant** //[[CL:Glossary:form]]// appears as a //[[CL:Glossary:top level form]]//, the //[[CL:Glossary:compiler]]// must recognize that //name// names a //[[CL:Glossary:constant variable]]//. An implementation may choose to evaluate the value-form at compile time, load time, or both. Therefore, users must ensure that the //initial-value// can be //[[CL:Glossary:evaluate|evaluated]]// at compile time (regardless of whether or not references to //name// appear in the file) and that it always //[[CL:Glossary:evaluate|evaluates]]// to the same value.

====Examples====
<blockquote>
(defconstant this-is-a-constant 'never-changing "for a test") <r>THIS-IS-A-CONSTANT </r>
this-is-a-constant <r>NEVER-CHANGING</r>
(documentation 'this-is-a-constant 'variable) <r>"for a test"</r>
(constantp 'this-is-a-constant) <r>//[[CL:Glossary:true]]//</r>
</blockquote>

====Affected By====
None.

====Exceptional Situations====
None.

====See Also====
  * **[[CL:Macros:declaim|Function DECLAIM]]**
  * **[[CL:Macros:defparameter|Macro DEFPARAMETER]]**
  * **[[CL:Macros:defvar|Macro DEFVAR]]**
  * **[[CL:Functions:documentation|Generic Function DOCUMENTATION]]**
  * **[[CL:Functions:proclaim|Function PROCLAIM]]**
  * {\secref\ConstantVars}
  * {\secref\Compilation}

====Notes====
None.

\issue{DEFVAR-DOCUMENTATION:UNEVALUATED} \issue{DEFINING-MACROS-NON-TOP-LEVEL:ALLOW} \issue{DEFCONSTANT-SPECIAL:NO} \issue{COMPILE-FILE-HANDLING-OF-TOP-LEVEL-FORMS:CLARIFY}
