====== Accessor FIND-CLASS ======

====Syntax====
**find-class {symbol** //\opt} errorp environment// → //class// 
(**setf** (**find-class {symbol** //\opt} errorp environment//) //new-class//)

====Arguments and Values====
//symbol// - a //[[CL:Glossary:symbol]]//.

//errorp// - a //[[CL:Glossary:generalized boolean]]//. The default is //[[CL:Glossary:true]]//.

//environment// -- same as the **''&environment''** argument to macro expansion functions and is used to distinguish between compile-time and run-time environments.

The **''&environment''** argument has //[[CL:Glossary:dynamic extent]]//; the consequences are undefined if the **''&environment''** argument is referred to outside the //[[CL:Glossary:dynamic extent]]// of the macro expansion function.

//class// - a //[[CL:Glossary:class]]// //[[CL:Glossary:object]]//, or **[[CL:Constant Variables:nil]]**.

====Description====
Returns the //[[CL:Glossary:class]]// //[[CL:Glossary:object]]// named by the //symbol// in the //environment//. If there is no such //[[CL:Glossary:class]]//, **[[CL:Constant Variables:nil]]** is returned if //errorp// is //[[CL:Glossary:false]]//; otherwise, if //errorp// is //[[CL:Glossary:true]]//, an error is signaled.

The //[[CL:Glossary:class]]// associated with a particular //[[CL:Glossary:symbol]]// can be changed by using **[[CL:Macros:setf]]** with **[[CL:Functions:find-class]]**;

or, if the new //[[CL:Glossary:class]]// given to **[[CL:Macros:setf]]** is **[[CL:Constant Variables:nil]]**, the //[[CL:Glossary:class]]// association is removed (but the //[[CL:Glossary:class]]// //[[CL:Glossary:object]]// itself is not affected).

The results are undefined if the user attempts to change

or remove

the //[[CL:Glossary:class]]// associated with a //[[CL:Glossary:symbol]]// that is defined as a //[[CL:Glossary:type specifier]]// in this standard. see section {\secref\IntegratingTypesAndClasses}.

When using \SETFof{find-class}, any //[[CL:Glossary:errorp]]// argument is //[[CL:Glossary:evaluate|evaluated]]// for effect, but any //[[CL:Glossary:values]]// it returns are ignored; the //errorp// //[[CL:Glossary:parameter]]// is permitted primarily so that the //environment// //[[CL:Glossary:parameter]]// can be used.

The //environment// might be used to distinguish between a compile-time and a run-time environment.

====Examples====
None.

====Affected By====
None.

====Exceptional Situations====
If there is no such //[[CL:Glossary:class]]// and //errorp// is //[[CL:Glossary:true]]//, **[[CL:Functions:find-class]]** signals an error of type **[[CL:Types:error]]**.

====See Also====
**[[CL:Macros:defmacro]]**, {\secref\IntegratingTypesAndClasses}

====Notes====
None.

\issue{MACRO-ENVIRONMENT-EXTENT:DYNAMIC} \issue{SETF-FIND-CLASS:ALLOW-NIL} \issue{SETF-FIND-CLASS:ALLOW-NIL}
