====== Macro TRACE, UNTRACE ======

====Syntax====

\DefmacWithValues trace {\starparam{function-name}} {trace-result} \DefmacWithValues untrace {\starparam{function-name}} {untrace-result}

====Arguments and Values====

//function-name// - a //[[CL:Glossary:function name]]//.

//trace-result// - //[[CL:Glossary:implementation-dependent]]//, unless no //[[CL:Glossary:function-names]]// are supplied,

in which case //trace-result// is a //[[CL:Glossary:list]]// of //[[CL:Glossary:function names]]//.

//untrace-result// - //[[CL:Glossary:implementation-dependent]]//.

====Description====

**[[CL:Macros:trace]]** and **[[CL:Macros:untrace]]** control the invocation of the trace facility.

Invoking **[[CL:Macros:trace]]** with one or more //function-names// causes the denoted //[[CL:Glossary:functions]]// to be "traced." Whenever a traced //[[CL:Glossary:function]]// is invoked, information about the call, about the arguments passed, and about any eventually returned values is printed to //[[CL:Glossary:trace output]]//. If **[[CL:Macros:trace]]** is used with no //function-names//, no tracing action is performed; instead, a list of the //[[CL:Glossary:functions]]// currently being traced is returned.

Invoking **[[CL:Macros:untrace]]** with one or more function names causes those functions to be "untraced" (i.e. no longer traced).

If **[[CL:Macros:untrace]]** is used with no //function-names//, all //[[CL:Glossary:functions]]// currently being traced are untraced.

If a //[[CL:Glossary:function]]// to be traced has been open-coded (//e.g.// because it was declared \declref{inline}), a call to that //[[CL:Glossary:function]]// might not produce trace output.

====Examples====

<blockquote> (defun fact (n) (if (zerop n) 1 (* n (fact (- n 1))))) → FACT (trace fact) → (FACT) ;; Of course, the format of traced output is implementation-dependent. (fact 3)
▷ 1 Enter FACT 3
▷ | 2 Enter FACT 2
▷ | 3 Enter FACT 1
▷ | | 4 Enter FACT 0
▷ | | 4 Exit FACT 1
▷ | 3 Exit FACT 1
▷ | 2 Exit FACT 2
▷ 1 Exit FACT 6 → 6 </blockquote>

====Side Effects====

Might change the definitions of the //[[CL:Glossary:functions]]// named by //function-names//.

====Affected By====

Whether the functions named are defined or already being traced.

====Exceptional Situations====

Tracing an already traced function, or untracing a function not currently being traced, should produce no harmful effects, but might signal a warning.

====See Also====

**[[CL:Variables:*trace-output*]]**, **[[CL:Macros:step]]**

====Notes====

**[[CL:Macros:trace]]** and **[[CL:Macros:untrace]]** may also accept additional //[[CL:Glossary:implementation-dependent]]// argument formats. The format of the trace output is //[[CL:Glossary:implementation-dependent]]//.

Although **[[CL:Macros:trace]]** can be extended to permit non-standard options, //[[CL:Glossary:implementations]]// are nevertheless encouraged (but not required) to warn about the use of syntax or options that are neither specified by this standard nor added as an extension by the //[[CL:Glossary:implementation]]//, since they could be symptomatic of typographical errors or of reliance on features supported in //[[CL:Glossary:implementations]]// other than the current //[[CL:Glossary:implementation]]//.

\issue{FUNCTION-NAME:LARGE}
