====== Function EXP, EXPT ======

====Syntax====
  * **exp** //number// → //result// 
  * **expt** //base-number power-number// → //result//

====Arguments and Values====
  * //number// - a //[[CL:Glossary:number]]//.
  * //base-number// - a //[[CL:Glossary:number]]//.
  * //power-number// - a //[[CL:Glossary:number]]//.
  * //result// - a //[[CL:Glossary:number]]//.

====Description====
**exp** and **expt** perform exponentiation.

**exp** returns ''e'' raised to the power //number//, where ''e'' is the base of the natural logarithms.

**exp** has no branch cut.

**expt** returns //base-number// raised to the power //power-number//. If the //base-number// is a //[[CL:Glossary:rational]]// and //power-number// is an //[[CL:Glossary:integer]]//, the calculation is exact and the result will be of type **[[CL:Types:rational]]**; otherwise a floating-point approximation might result.

For **expt** of a //[[CL:Glossary:complex rational]]// to an //[[CL:Glossary:integer]]// power, the calculation must be exact and the result is of type ''([[CL:Types:or]] [[CL:Types:rational]] ([[CL:Types:complex]] [[CL:Types:rational]]))''.

The result of **expt** can be a //[[CL:Glossary:complex]]//, even when neither argument is a //[[CL:Glossary:complex]]//, if //base-number// is negative and //power-number// is not an //[[CL:Glossary:integer]]//. The result is always the //[[CL:Glossary:principal]]// //[[CL:Glossary:complex]]// //[[CL:Glossary:value]]//. For example, ''(expt -8 1/3)'' is not permitted to return ''-2'', even though ''-2'' is one of the cube roots of ''-8''. The //[[CL:Glossary:principal]]// cube root is a //[[CL:Glossary:complex]]// approximately equal to ''#C(1.0 1.73205)'', not ''-2''.

**expt** is defined as ''b<sup>x</sup> = ''e<sup>x log b</sup>''. This defines the //[[CL:Glossary:principal]]// //[[CL:Glossary:values]]// precisely. The range of **expt** is the entire complex plane. Regarded as a function of //x//, with //b// fixed, there is no branch cut. Regarded as a function of //b//, with //x// fixed, there is in general a branch cut along the negative real axis, continuous with quadrant II. The domain excludes the origin. By definition, ''0<sup>0</sup> = 1''. If ''//b// = 0'' and the real part of //x// is strictly positive, then ''b<sup>x</sup> = 0''. For all other values of //x//, ''0<sup>x</sup>'' is an error.

When //power-number// is an //[[CL:Glossary:integer]]// ''0'', then the result is always the value one in the //[[CL:Glossary:type]]// of //base-number//, even if the //base-number// is zero (of any //[[CL:Glossary:type]]//). That is:

<blockquote>
(expt x 0) ≡ ([[CL:Functions:coerce]] 1 ([[CL:Functions:type-of]] x))
</blockquote> 

If //power-number// is a zero of any other //[[CL:Glossary:type]]//, then the result is also the value one, in the //[[CL:Glossary:type]]// of the arguments after the application of the contagion rules in \secref\NumericContagionRules, with one exception: the consequences are undefined if //base-number// is zero when //power-number// is zero and not of type **[[CL:Types:integer]]**.

====Examples====

<blockquote>
(exp 0) → 1.0 
(exp 1) → 2.718282 
(exp ([[CL:Functions:log]] 5)) → 5.0 
(expt 2 8) → 256 
(expt 4 .5) → 2.0 
(expt #c(0 1) 2) → -1 
(expt #c(2 2) 3) → #C(-16 16) 
(expt #c(2 2) 4) → -64 
</blockquote>

====Affected By====
None.

====Exceptional Situations====
None.

====See Also====
  * **[[CL:Functions:log|Function LOG]]**
  * {\secref\FloatSubstitutability}

====Notes====
Implementations of **expt** are permitted to use different algorithms for the cases of a //power-number// of type **[[CL:Types:rational]]** and a //power-number// of type **[[CL:Types:float]]**.

Note that by the following logic, ''(sqrt (expt //x// 3))'' is not equivalent to ''(expt //x// 3/2)''.

<blockquote>
([[CL:Macros:defparameter]] *x* (exp (/ (* 2 pi #c(0 1)) 3))) ; e<sup>2πi/3</sup>
(expt *x* 3) → 1 ; except for round-off error 
(sqrt (expt *x* 3)) → 1 ; except for round-off error 
(expt *x* 3/2) → -1 ; except for round-off error </blockquote>

\issue{COMPLEX-RATIONAL-RESULT:EXTEND} \issue{COMPLEX-RATIONAL-RESULT:EXTEND} \issue{EXPT-RATIO:P.211}
