====== Special Operator FUNCTION ======

====Syntax====
  * **function** //name// → //function//

====Arguments and Values====
  * //name// - a //[[CL:Glossary:function name]]// or //[[CL:Glossary:lambda expression]]//.
  * //function// - a //[[CL:Glossary:function]]// //[[CL:Glossary:object]]//.

====Description====
The //[[CL:Glossary:value]]// of **function** is the //[[CL:Glossary:functional value]]// of //name// in the current //[[CL:Glossary:lexical environment]]//.

If //name// is a //[[CL:Glossary:function name]]//, the functional definition of that name is that established by the innermost lexically enclosing **[[CL:Special Operators:flet]]**, **[[CL:Special Operators:labels]]**, or **[[CL:Special Operators:macrolet]]** //[[CL:Glossary:form]]//, if there is one. Otherwise the global functional definition of the //[[CL:Glossary:function name]]// is returned.

If //name// is a //[[CL:Glossary:lambda expression]]//, then a //[[CL:Glossary:lexical closure]]// is returned. In situations where a //[[CL:Glossary:closure]]// over the same set of //[[CL:Glossary:binding|bindings]]// might be produced more than once, the various resulting //[[CL:Glossary:closures]]// might or might not be **[[CL:Functions:eq]]**.

It is an error to use **function** on a //[[CL:Glossary:function name]]// that does not denote a //[[CL:Glossary:function]]// in the lexical environment in which the **function** form appears. Specifically, it is an error to use **function** on a //[[CL:Glossary:symbol]]// that denotes a //[[CL:Glossary:macro]]// or //[[CL:Glossary:special form]]//. An implementation may choose not to signal this error for performance reasons, but implementations are forbidden from defining the failure to signal an error as a useful behavior.

====Examples====
<blockquote>
(defun adder (x)
  (function (lambda (y) (+ x y)))) <r>ADDER</r>
</blockquote>

The result of ''(adder 3)'' is a function that adds ''3'' to its argument:

<blockquote>
([[CL:Macros:defparameter]] *add3* (adder 3)) <r>ADD3</r>
(funcall *add3* 5) <r>8</r>
</blockquote>

This works because **function** creates a //[[CL:Glossary:closure]]// of the //[[CL:Glossary:lambda expression]]// that is able to refer to the //[[CL:Glossary:value]]// ''3'' of the variable ''x'' even after control has returned from the function ''adder''.

====Side Effects====
None.

====Affected By====
None.

====Exceptional Situations====
None.

====See Also====
  * **[[CL:Macros:defun|Macro DEFUN]]**
  * **[[CL:Functions:fdefinition|Function FDEFINITION]]**
  * **[[CL:Special Operators:flet|Special Operator FLET]]**
  * **[[CL:Special Operators:labels|Special Operator LABELS]]**
  * **[[CL:Functions:symbol-function|Function SYMBOL-FUNCTION]]**
  * {\secref\SymbolsAsForms}
  * {\secref\SharpsignQuote}
  * {\secref\PrintingOtherObjects}

====Notes====
The notation ''#'//name//'' may be used as an abbreviation for ''(function //name//)''.

\issue{FUNCTION-NAME:LARGE} \issue{FUNCTION-TYPE:X3J13-MARCH-88}
