====== Function CONCATENATE ======

====Syntax====

**concatenate {result-type** //\rest} sequences// → //result-sequence//

====Arguments and Values====

//result-type// - a **[[CL:Types:sequence]]** //[[CL:Glossary:type specifier]]//.

//sequences// - a //[[CL:Glossary:sequence]]//.

//result-sequence// - a //[[CL:Glossary:proper sequence]]// of //[[CL:Glossary:type]]// //result-type//.

====Description====

**[[CL:Functions:concatenate]]** returns a //[[CL:Glossary:sequence]]// that contains all the individual elements of all the //sequences// in the order that they are supplied. The //[[CL:Glossary:sequence]]// is of type //result-type//, which must be a subtype of **[[CL:Types:sequence]]**.

All of the //sequences// are copied from; the result does not share any structure with any of the //sequences//.

Therefore, if only one //sequence// is provided and it is of type //result-type//, **[[CL:Functions:concatenate]]** is required to copy //sequence// rather than simply returning it.

It is an error if any element of the //sequences// cannot be an element of the //[[CL:Glossary:sequence]]// result. \reviewer{Barmar: Should signal?}

If the //result-type// is a //[[CL:Glossary:subtype]]// of **[[CL:Types:list]]**,the result will be a //[[CL:Glossary:list]]//.

If the //result-type// is a //[[CL:Glossary:subtype]]// of **[[CL:Types:vector]]**,then if the implementation can determine the element type specified for the //result-type//, the element type of the resulting array is the result of //[[CL:Glossary:upgrade|upgrading]]// that element type; or, if the implementation can determine that the element type is unspecified (or ''*''), the element type of the resulting array is **[[CL:Types:t]]**; otherwise, an error is signaled.

====Examples====

<blockquote> (concatenate 'string "all" " " "together" " " "now") → "all together now" (concatenate 'list "ABC" '(d e f) #(1 2 3) #*1011) → (#\\A #\\B #\\C D E F 1 2 3 1 0 1 1) (concatenate 'list) → NIL </blockquote>

<blockquote> (concatenate '(vector * 2) "a" "bc") should signal an error </blockquote>

====Affected By====

None.

====Exceptional Situations====

An error is signaled if the //result-type// is neither a //[[CL:Glossary:recognizable subtype]]// of **[[CL:Types:list]]**, nor a //[[CL:Glossary:recognizable subtype]]// of **[[CL:Types:vector]]**.

An error of type **[[CL:Types:type-error]]** should be signaled if //result-type// specifies the number of elements and the sum of //sequences// is different from that number.

====See Also====

**[[CL:Functions:append]]**

====Notes====

None.



\issue{SEQUENCE-TYPE-LENGTH:MUST-MATCH} \issue{CONCATENATE-SEQUENCE:SIGNAL-ERROR} \issue{CONCATENATE-SEQUENCE:SIGNAL-ERROR} \issue{SEQUENCE-TYPE-LENGTH:MUST-MATCH} \issue{CONCATENATE-SEQUENCE:SIGNAL-ERROR}
