====== Function INVOKE-RESTART-INTERACTIVELY ======

====Syntax====

\DefunWithValues invoke-restart-interactively {restart} {\starparam{result}}

====Arguments and Values====

//restart// - a //[[CL:Glossary:restart designator]]//.

//results// - the //[[CL:Glossary:values]]// returned by the //[[CL:Glossary:function]]// associated with //restart//, if that //[[CL:Glossary:function]]// returns.

====Description====

**[[CL:Functions:invoke-restart-interactively]]** calls the //[[CL:Glossary:function]]// associated with //restart//, prompting for any necessary arguments. If //restart// is a name, it must be valid in the current //[[CL:Glossary:dynamic environment]]//.

**[[CL:Functions:invoke-restart-interactively]]** prompts for arguments by executing the code provided in the **'':interactive''** keyword to **[[CL:Macros:restart-case]]** or **'':interactive-function''** keyword to **[[CL:Macros:restart-bind]]**.

If no such options have been supplied in the corresponding **[[CL:Macros:restart-bind]]** or **[[CL:Macros:restart-case]]**, then the consequences are undefined if the //restart// takes required arguments. If the arguments are optional, an argument list of **[[CL:Constant Variables:nil]]** is used.

Once the arguments have been determined, **[[CL:Functions:invoke-restart-interactively]]** executes the following:

<blockquote> (apply #'invoke-restart ''restart'' ''arguments'') </blockquote>


====Examples====

<blockquote> (defun add3 (x) (check-type x number) (+ x 3))

(add3 'seven)
▷ Error: The value SEVEN was not of type NUMBER.
▷ To continue, type :CONTINUE followed by an option number:
▷ 1: Specify a different value to use.
▷ 2: Return to Lisp Toplevel.
▷ Debug> \IN{(invoke-restart-interactively 'store-value)}
▷ Type a form to evaluate and use: \IN{7} → 10 </blockquote>

====Side Effects====

If prompting for arguments is necesary, some typeout may occur (on //[[CL:Glossary:query I/O]]//).


A non-local transfer of control might be done by the restart.

====Affected By====

**[[CL:Variables:*query-io*]]**, active //[[CL:Glossary:restarts]]//

====Exceptional Situations====

If //restart// is not valid, an error of type **[[CL:Types:control-error]]** is signaled.

====See Also====

**[[CL:Functions:find-restart]]**, **[[CL:Functions:invoke-restart]]**, **[[CL:Macros:restart-case]]**, **[[CL:Macros:restart-bind]]**

====Notes====

**[[CL:Functions:invoke-restart-interactively]]** is used internally by the debugger and may also be useful in implementing other portable, interactive debugging tools.

