====== Standard Generic Function SLOT-MISSING ======

====Syntax====

\DefgenWithValues slot-missing {class object slot-name operation ''&optional'' new-value} {\starparam{result}}

====Method Signatures====

\Defmeth slot-missing {\vtop{\hbox{\specparam{class}{t} object slot-name} operation ''&optional'' new-value}}

====Arguments and Values====

//class// - the //[[CL:Glossary:class]]// of //object//.

//object// - an //[[CL:Glossary:object]]//.

//slot-name// - a //[[CL:Glossary:symbol]]// (the //[[CL:Glossary:name]]// of a would-be //[[CL:Glossary:slot]]//).

//operation// - one of the //[[CL:Glossary:symbols]]// **[[CL:Functions:setf]]**, **[[CL:Functions:slot-boundp]]**, **[[CL:Functions:slot-makunbound]]**, or **[[CL:Functions:slot-value]]**.

//new-value// - an //[[CL:Glossary:object]]//.

//result// - an //[[CL:Glossary:object]]//.

====Description====

The generic function **[[CL:Functions:slot-missing]]** is invoked when an attempt is made to //[[CL:Glossary:access]]// a //[[CL:Glossary:slot]]// in an //object// whose //[[CL:Glossary:metaclass]]// is **[[CL:Types:standard-class]]** and the //[[CL:Glossary:slot]]// of the name //slot-name// is not a //[[CL:Glossary:name]]// of a //[[CL:Glossary:slot]]// in that //[[CL:Glossary:class]]//. The default //[[CL:Glossary:method]]// signals an error.

The generic function **[[CL:Functions:slot-missing]]** is not intended to be called by programmers. Programmers may write //[[CL:Glossary:methods]]// for it.

The generic function **[[CL:Functions:slot-missing]]** may be called during evaluation of **[[CL:Functions:slot-value]]**, ''(setf slot-value)'', **[[CL:Functions:slot-boundp]]**, and **[[CL:Functions:slot-makunbound]]**. For each of these operations the corresponding //[[CL:Glossary:symbol]]// for the //operation// argument is \misc{slot-value}, \misc{setf}, \misc{slot-boundp}, and \misc{slot-makunbound} respectively.

The optional //new-value// argument to **[[CL:Functions:slot-missing]]** is used when the operation is attempting to set the value of the //[[CL:Glossary:slot]]//.


If **[[CL:Functions:slot-missing]]** returns, its values will be treated as follows:

\beginlist \item{\bull} If the //operation// is \misc{setf} or \misc{slot-makunbound}, any //[[CL:Glossary:values]]// will be ignored by the caller.

\item{\bull} If the //operation// is \misc{slot-value}, only the //[[CL:Glossary:primary value]]// will be used by the caller, and all other values will be ignored.

\item{\bull} If the //operation// is \misc{slot-boundp}, any //[[CL:Glossary:boolean equivalent]]// of the //[[CL:Glossary:primary value]]// of the //[[CL:Glossary:method]]// might be is used, and all other values will be ignored. \endlist

====Examples====

None.

====Affected By====

None.

====Exceptional Situations====

The default //[[CL:Glossary:method]]// on **[[CL:Functions:slot-missing]]** signals an error of type **[[CL:Types:error]]**.

====See Also====

**[[CL:Macros:defclass]]**, **[[CL:Functions:slot-exists-p]]**, **[[CL:Functions:slot-value]]**

====Notes====

The set of arguments (including the //[[CL:Glossary:class]]// of the instance) facilitates defining methods on the metaclass for **[[CL:Functions:slot-missing]]**.

\issue{SLOT-MISSING-VALUES:SPECIFY}
