====== Function COMPLEX ======

====Syntax====

**complex {realpart** //\opt} imagpart// → //complex//

====Arguments and Values====

//realpart// - a //[[CL:Glossary:real]]//.

//imagpart// - a //[[CL:Glossary:real]]//.

//complex// - a //[[CL:Glossary:rational]]// or a //[[CL:Glossary:complex]]//.

====Description====

**[[CL:Functions:complex]]** returns a //[[CL:Glossary:number]]// whose real part is //realpart// and whose imaginary part is //imagpart//.

If //realpart// is a //[[CL:Glossary:rational]]// and //imagpart// is the //[[CL:Glossary:rational]]// number zero, the result of **[[CL:Functions:complex]]** is //realpart//, a //[[CL:Glossary:rational]]//. Otherwise, the result is a //[[CL:Glossary:complex]]//.

If either //realpart// or //imagpart// is a //[[CL:Glossary:float]]//, the non-//[[CL:Glossary:float]]// is converted to a //[[CL:Glossary:float]]// before the //[[CL:Glossary:complex]]// is created. If //imagpart// is not supplied, the imaginary part is a zero of the same //[[CL:Glossary:type]]// as //realpart//; i.e. **[[CL:Functions:(coerce 0 (type-of //realpart//))]]** is effectively used.

Type upgrading implies a movement upwards in the type hierarchy lattice.

In the case of //[[CL:Glossary:complexes]]//, the //type-specifier// \reviewer{Barmar: What type specifier?} must be a subtype of **[[CL:Functions:(upgraded-complex-part-type //type-specifier//)]]**. If //type-specifier1// is a subtype of //type-specifier2//, then **[[CL:Functions:(upgraded-complex-element-type '//type-specifier1//)]]** must also be a subtype of **[[CL:Functions:(upgraded-complex-element-type '//type-specifier2//)]]**. Two disjoint types can be upgraded into the same thing.


====Examples==== <blockquote> (complex 0) → 0 (complex 0.0) → #C(0.0 0.0) (complex 1 1/2) → #C(1 1/2) (complex 1 .99) → #C(1.0 0.99) (complex 3/2 0.0) → #C(1.5 0.0) </blockquote>

====Side Effects====

None.

====Affected By====

None.

====Exceptional Situations====

None.

====See Also====

**[[CL:Functions:realpart]]**, **[[CL:Functions:imagpart]]**,

{\secref\SharpsignC}

====Notes====

None.

\issue{ARRAY-TYPE-ELEMENT-TYPE-SEMANTICS:UNIFY-UPGRADING}
