

\issue{EXIT-EXTENT:MINIMAL} When a transfer of control is initiated by \specref{go},  \specref{return-from}, or \specref{throw} the following events occur in order to accomplish the transfer of control. Note that for \specref{go},  the //[[CL:Glossary:exit point]]// is the //[[CL:Glossary:form]]// within the \specref{tagbody} that is being executed at the time the \specref{go} is performed; for \specref{return-from}, the //[[CL:Glossary:exit point]]// is the corresponding  \specref{block} //[[CL:Glossary:form]]//; and for \specref{throw}, the //[[CL:Glossary:exit point]]// is the corresponding  \specref{catch} //[[CL:Glossary:form]]//. 

\beginlist \itemitem{1.} 
 Intervening //[[CL:Glossary:exit points]]// are ``abandoned''
 (\ie their //[[CL:Glossary:extent]]// ends 
      and it is no longer valid to attempt to transfer control through them).

\itemitem{2.} 
 The cleanup clauses of any intervening \specref{unwind-protect} clauses
 are evaluated.
  \itemitem{3.} 

 Intervening dynamic //[[CL:Glossary:bindings]]// of **[[CL:Declarations:special]]** variables,
 //[[CL:Glossary:catch tags]]//, //[[CL:Glossary:condition handlers]]//, and //[[CL:Glossary:restarts]]//
 are undone.
  \itemitem{4.} 
 The //[[CL:Glossary:extent]]// of the //[[CL:Glossary:exit point]]// being invoked ends,
 and control is passed to the target. \endlist 
  The extent of an exit being ``abandoned'' because it is being passed over ends as soon as the transfer of control is initiated. That is, event 1 occurs at the beginning of the initiation of the transfer of control.  The consequences are undefined if an attempt is made to transfer control  to an //[[CL:Glossary:exit point]]// whose //[[CL:Glossary:dynamic extent]]// has ended.
 

Events 2 and 3 are actually performed interleaved, in the order corresponding to the reverse order in which they were established. The effect of this is that the cleanup clauses of an \specref{unwind-protect} see the same dynamic //[[CL:Glossary:bindings]]//  of variables and //[[CL:Glossary:catch tags]]// as were visible when the \specref{unwind-protect} was entered.
  Event 4 occurs at the end of the transfer of control.
 

