



This section describes the algorithm used by the //[[CL:Glossary:Lisp reader]]//
to parse //[[CL:Glossary:objects]]// from an //[[CL:Glossary:input]]// //[[CL:Glossary:character]]// //[[CL:Glossary:stream]]//,
including how the //[[CL:Glossary:Lisp reader]]// processes //[[CL:Glossary:macro characters]]//.

When dealing with //[[CL:Glossary:tokens]]//, the reader's basic function is to distinguish
representations of //[[CL:Glossary:symbols]]// from those of //[[CL:Glossary:numbers]]//.




When a //[[CL:Glossary:token]]// is accumulated, it is assumed to represent a //[[CL:Glossary:number]]// if it
satisfies the syntax for numbers listed in \figref\SyntaxForNumericTokens.




If it does not represent a //[[CL:Glossary:number]]//,
it is then assumed to be a //[[CL:Glossary:potential number]]// 
if it satisfies the rules governing the syntax for a //[[CL:Glossary:potential number]]//.
If a valid //[[CL:Glossary:token]]// is neither a representation of a //[[CL:Glossary:number]]// 
			       nor a //[[CL:Glossary:potential number]]//,
it represents a //[[CL:Glossary:symbol]]//.

The algorithm performed by the //[[CL:Glossary:Lisp reader]]// is as follows:


\beginlist
\item{1.}            
If at end of file, end-of-file processing is performed as specified
in **[[CL:Functions:read]]**.
Otherwise,
one //[[CL:Glossary:character]]//, //x//,  is read from the //[[CL:Glossary:input]]// //[[CL:Glossary:stream]]//, and
dispatched according to the //[[CL:Glossary:syntax type]]// of //x// to one
of steps 2 to 7.


\item{2.}                                          
If //x// is an //[[CL:Glossary:invalid]]// //[[CL:Glossary:character]]//,
an error \oftype{reader-error} is signaled.


\item{3.}
If //x// is a //[[CL:Glossary:whitespace]]//\meaning{2} //[[CL:Glossary:character]]//,
then it is discarded and step 1 is re-entered.


\item{4.}
If //x// is a //[[CL:Glossary:terminating]]// or //[[CL:Glossary:non-terminating]]// //[[CL:Glossary:macro character]]//
then its associated //[[CL:Glossary:reader macro function]]// is called with two //[[CL:Glossary:arguments]]//,
the //[[CL:Glossary:input]]// //[[CL:Glossary:stream]]// and //x//.


The //[[CL:Glossary:reader macro function]]// may read //[[CL:Glossary:characters]]// 
from the //[[CL:Glossary:input]]// //[[CL:Glossary:stream]]//; 
if it does, it will see those //[[CL:Glossary:characters]]// following the //[[CL:Glossary:macro character]]//.
The //[[CL:Glossary:Lisp reader]]// may be invoked recursively from the //[[CL:Glossary:reader macro function]]//.


The //[[CL:Glossary:reader macro function]]// must not have any side effects other than on the
//[[CL:Glossary:input]]// //[[CL:Glossary:stream]]//;
because of backtracking and restarting of the **[[CL:Functions:read]]** operation,
front ends to the //[[CL:Glossary:Lisp reader]]// (\eg ``editors'' and ``rubout handlers'') 
may cause the //[[CL:Glossary:reader macro function]]// to be called repeatedly during the
reading of a single //[[CL:Glossary:expression]]// in which //x// only appears once.


The //[[CL:Glossary:reader macro function]]// may return zero values or one value.
If one value is returned,
then that value is returned as the result of the read operation;
the algorithm is done.
If zero values are returned, then step 1 is re-entered.


\item{5.}
If //x// is a //[[CL:Glossary:single escape]]// //[[CL:Glossary:character]]//
then the next //[[CL:Glossary:character]]//, //y//, is read, or an error \oftype{end-of-file} 
is signaled if at the end of file.
//y// is treated as if it is a //[[CL:Glossary:constituent]]// 
whose only //[[CL:Glossary:constituent trait]]// is //[[CL:Glossary:alphabetic]]//\meaning{2}.
//y// is used to begin a //[[CL:Glossary:token]]//, and step 8 is entered.


\item{6.}
If //x// is a //[[CL:Glossary:multiple escape]]// //[[CL:Glossary:character]]//
then a //[[CL:Glossary:token]]// (initially
containing no //[[CL:Glossary:characters]]//) is  begun and step 9 is entered.


\item{7.}
If //x// is a //[[CL:Glossary:constituent]]// //[[CL:Glossary:character]]//, then it begins a //[[CL:Glossary:token]]//.
After the //[[CL:Glossary:token]]// is read in, it will be interpreted
either as a \Lisp\ //[[CL:Glossary:object]]// or as being of invalid syntax.
If the //[[CL:Glossary:token]]// represents an //[[CL:Glossary:object]]//,
that //[[CL:Glossary:object]]// is returned as the result of the read operation.
If the //[[CL:Glossary:token]]// is of invalid syntax, an error is signaled.



If //x// is a //[[CL:Glossary:character]]// with //[[CL:Glossary:case]]//,
it might be replaced with the corresponding //[[CL:Glossary:character]]// of the opposite //[[CL:Glossary:case]]//, 
depending on the //[[CL:Glossary:readtable case]]// of the //[[CL:Glossary:current readtable]]//,
as outlined in \secref\ReadtableCaseReadEffect.
//X// is used to begin a //[[CL:Glossary:token]]//, and step 8 is entered.




\item{8.}
At this point a //[[CL:Glossary:token]]// is being accumulated, and an even number
of //[[CL:Glossary:multiple escape]]// //[[CL:Glossary:characters]]// have been encountered.
If at end of file, step 10 is entered.
Otherwise, a //[[CL:Glossary:character]]//, //y//, is read, and
one of the following actions is performed according to its //[[CL:Glossary:syntax type]]//:

\beginlist
\itemitem{\bull}
If //y// is a //[[CL:Glossary:constituent]]// or //[[CL:Glossary:non-terminating]]// //[[CL:Glossary:macro character]]//:
\beginlist
\itemitem{--}



If //y// is a //[[CL:Glossary:character]]// with //[[CL:Glossary:case]]//,
it might be replaced with the corresponding //[[CL:Glossary:character]]// of the opposite //[[CL:Glossary:case]]//, 
depending on the //[[CL:Glossary:readtable case]]// of the //[[CL:Glossary:current readtable]]//,
as outlined in \secref\ReadtableCaseReadEffect.
\itemitem{--}
//Y// is appended to the //[[CL:Glossary:token]]// being built.
\itemitem{--}
Step 8 is repeated.
\endlist


\itemitem{\bull}
If //y// is a //[[CL:Glossary:single escape]]// //[[CL:Glossary:character]]//, then the next //[[CL:Glossary:character]]//,
//z//, is read, or an error \oftype{end-of-file} is signaled if at end of file.
//Z// is treated as if it is a //[[CL:Glossary:constituent]]// 
whose only //[[CL:Glossary:constituent trait]]// is //[[CL:Glossary:alphabetic]]//\meaning{2}.
//Z// is appended to the //[[CL:Glossary:token]]// being built,
and step 8 is repeated.


\itemitem{\bull}
If //y// is a //[[CL:Glossary:multiple escape]]// //[[CL:Glossary:character]]//,
then step 9 is entered.


\itemitem{\bull}
If //y// is an //[[CL:Glossary:invalid]]// //[[CL:Glossary:character]]//,
an error \oftype{reader-error} is signaled.


\itemitem{\bull}
If //y// is a //[[CL:Glossary:terminating]]// //[[CL:Glossary:macro character]]//,
then it terminates the //[[CL:Glossary:token]]//.
First the //[[CL:Glossary:character]]// //y// is unread (see **[[CL:Functions:unread-char]]**),
and then step 10 is entered.


\itemitem{\bull}
If //y// is a //[[CL:Glossary:whitespace]]//\meaning{2} //[[CL:Glossary:character]]//, then it terminates
the //[[CL:Glossary:token]]//.  First the //[[CL:Glossary:character]]// //y// is unread
if appropriate (see **[[CL:Functions:read-preserving-whitespace]]**),
and then step 10 is entered.
\endlist



\item{9.}
At this point a //[[CL:Glossary:token]]// is being accumulated, and an odd number
of //[[CL:Glossary:multiple escape]]// //[[CL:Glossary:characters]]// have been encountered.
If at end of file, an error \oftype{end-of-file} is signaled.
Otherwise, a //[[CL:Glossary:character]]//, //y//, is read, and
one of the following actions is performed according to its //[[CL:Glossary:syntax type]]//:


\beginlist
\itemitem{\bull}
If //y// is a //[[CL:Glossary:constituent]]//, macro, or //[[CL:Glossary:whitespace]]//\meaning{2} //[[CL:Glossary:character]]//,
//y// is treated as a //[[CL:Glossary:constituent]]// 
whose only //[[CL:Glossary:constituent trait]]// is //[[CL:Glossary:alphabetic]]//\meaning{2}.             
//Y// is appended to the //[[CL:Glossary:token]]// being built, and step 9 is repeated.


\itemitem{\bull}
If //y// is a //[[CL:Glossary:single escape]]// //[[CL:Glossary:character]]//, then the next //[[CL:Glossary:character]]//,
//z//, is read, or an error \oftype{end-of-file} is signaled if at end of file.
//Z// is treated as a //[[CL:Glossary:constituent]]//
whose only //[[CL:Glossary:constituent trait]]// is //[[CL:Glossary:alphabetic]]//\meaning{2}.
//Z// is appended to the //[[CL:Glossary:token]]// being built,
and step 9 is repeated.


\itemitem{\bull}
If //y// is a //[[CL:Glossary:multiple escape]]// //[[CL:Glossary:character]]//,
then step 8 is entered.


\itemitem{\bull}
If //y// is an //[[CL:Glossary:invalid]]// //[[CL:Glossary:character]]//,
an error \oftype{reader-error} is signaled.
\endlist


\item{10.}
An entire //[[CL:Glossary:token]]// has been accumulated.
The //[[CL:Glossary:object]]// represented by the //[[CL:Glossary:token]]// is returned 
as the result of the read operation,
or an error \oftype{reader-error} is signaled if the //[[CL:Glossary:token]]// is not of valid syntax.
\endlist











