
              
\Thefunction{change-class} can be used to change the //[[CL:Glossary:class]]// 
of an //[[CL:Glossary:instance]]// from its current class, $C\sub {\hbox{{\prmseven from}}}$,
to a different class, $C\sub {\hbox{{\prmseven to}}}$; it changes the
structure of the //[[CL:Glossary:instance]]// to conform to the definition of the class
$C\sub {\hbox{{\prmseven to}}}$.

Note that changing the //[[CL:Glossary:class]]// of an //[[CL:Glossary:instance]]// may cause
//[[CL:Glossary:slots]]// to be added or deleted.  Changing the //[[CL:Glossary:class]]// of an
//[[CL:Glossary:instance]]// does not change its identity as defined by the
**[[CL:Functions:eq]]** function.
      
When **[[CL:Functions:change-class]]** is invoked on an //[[CL:Glossary:instance]]//, a two-step
updating process takes place.  The first step modifies the structure of
the //[[CL:Glossary:instance]]// by adding new //[[CL:Glossary:local slots]]// and discarding 
//[[CL:Glossary:local slots]]// that are not specified in the new version of the //[[CL:Glossary:instance]]//.
The second step initializes the newly added //[[CL:Glossary:local slots]]// and performs 
any other user-defined actions. These two steps are further described in the 
two following sections.

\beginsubsection{Modifying the Structure of the Instance}

In order to make the //[[CL:Glossary:instance]]// conform to the class $C\sub
{\hbox{{\prmseven to}}}$, //[[CL:Glossary:local slots]]// specified by the class $C\sub
{\hbox{{\prmseven to}}}$ that are not specified by the class $C\sub
{\hbox{{\prmseven from}}}$ are added, and //[[CL:Glossary:local slots]]// not specified by
the class $C\sub {\hbox{{\prmseven to}}}$ that are specified by the
class $C\sub {\hbox{{\prmseven from}}}$ are discarded.

The values of //[[CL:Glossary:local slots]]// specified by both the class $C\sub
{\hbox{{\prmseven to}}}$ and the class $C\sub {\hbox{{\prmseven
from}}}$ are retained. If such a //[[CL:Glossary:local slot]]// was unbound, it remains
unbound.

The values of //[[CL:Glossary:slots]]// specified as shared in the class $C\sub
{\hbox{{\prmseven from}}}$ and as local in the class $C\sub
{\hbox{{\prmseven to}}}$ are retained.

This first step of the update does not affect the values of any 
//[[CL:Glossary:shared slots]]//.

\endsubsection%{Modifying the Structure of the Instance}

\beginsubsection{Initializing Newly Added Local Slots}
\DefineSection{InitNewLocalSlots}

The second step of the update initializes the newly added //[[CL:Glossary:slots]]// and
performs any other user-defined actions.  This step is implemented by
the generic function **[[CL:Functions:update-instance-for-different-class]]**.  The
generic function **[[CL:Functions:update-instance-for-different-class]]** is invoked
by **[[CL:Functions:change-class]]** after the first step of the update has been
completed.

\issue{CHANGE-CLASS-INITARGS:PERMIT}
The generic function **[[CL:Functions:update-instance-for-different-class]]** is
invoked on arguments computed by **[[CL:Functions:change-class]]**.
The first argument passed is a copy of the //[[CL:Glossary:instance]]// being updated 
and is an //[[CL:Glossary:instance]]// of the class $C\sub {\hbox{{\prmseven from}}}$; 
this copy has //[[CL:Glossary:dynamic extent]]// within the generic function **[[CL:Functions:change-class]]**.  
The second argument is the //[[CL:Glossary:instance]]// as updated so far by **[[CL:Functions:change-class]]**
and is an //[[CL:Glossary:instance]]// of the class $C\sub {\hbox{{\prmseven to}}}$.
The remaining arguments are an //[[CL:Glossary:initialization argument list]]//.






There is a system-supplied primary //[[CL:Glossary:method]]// for 
**[[CL:Functions:update-instance-for-different-class]]** that has two parameter
specializers, each of which is \theclass{standard-object}.  First
this //[[CL:Glossary:method]]// checks the validity of initialization arguments and
signals an error if an initialization argument is supplied that is not
declared as valid.  (For more information, \seesection\DeclaringInitargValidity.)
Then it calls the
generic function **[[CL:Functions:shared-initialize]]** with the following arguments:
the

new
//[[CL:Glossary:instance]]//, a list of //[[CL:Glossary:names]]// of the newly added 
//[[CL:Glossary:slots]]//, and the
initialization arguments it received.

\endsubsection%{Initializing Newly added Local Slots}

\beginsubsection{Customizing the Change of Class of an Instance}
             
//[[CL:Glossary:Methods]]// for **[[CL:Functions:update-instance-for-different-class]]** may be defined
to specify actions to be taken when an //[[CL:Glossary:instance]]// is updated.  If only
//[[CL:Glossary:after methods]]// for **[[CL:Functions:update-instance-for-different-class]]** are
defined, they will be run after the system-supplied primary //[[CL:Glossary:method]]// for
initialization and will not interfere with the default behavior of
**[[CL:Functions:update-instance-for-different-class]]**.









//[[CL:Glossary:Methods]]// 
for **[[CL:Functions:shared-initialize]]** may be defined to customize //[[CL:Glossary:class]]//
redefinition.  For more information, \seesection\SharedInitialize.
  
\endsubsection%{Customizing the Change of Class of an Instance}

