

Deprecated language features are not expected to appear in future \clisp\ standards, but are required to be implemented for conformance with this standard; \seesection\ReqLangFeatures.

//[[CL:Glossary:Conforming programs]]// can use deprecated features; however, it is considered good programming style to avoid them. It is permissible for the compiler to produce //[[CL:Glossary:style warnings]]//  about the use of such features at compile time,  but there should be no such warnings at program execution time.

\beginsubsection{Deprecated Functions}

\issue{REQUIRE-PATHNAME-DEFAULTS-AGAIN:X3J13-DEC-91} \issue{TEST-NOT-IF-NOT:FLUSH-ALL} \issue{GENTEMP-BAD-IDEA:DEPRECATE} The //[[CL:Glossary:functions]]// in \thenextfigure\ are deprecated. \displaythree{Deprecated Functions}{ assoc-if-not&nsubst-if-not&require\cr count-if-not&nsubstitute-if-not&set\cr delete-if-not&position-if-not&subst-if-not\cr find-if-not&provide&substitute-if-not\cr gentemp&rassoc-if-not&\cr member-if-not&remove-if-not&\cr }

\endsubsection%{Deprecated Functions}

\beginsubsection{Deprecated Argument Conventions}

\issue{GENSYM-NAME-STICKINESS:LIKE-TEFLON} The ability to pass a numeric //[[CL:Glossary:argument]]// to **[[CL:Functions:gensym]]** has been deprecated.

\issue{TEST-NOT-IF-NOT:FLUSH-ALL} \Thekeyarg{test-not} to the //[[CL:Glossary:functions]]// in \thenextfigure\ are deprecated.

\displaythree{Functions with Deprecated :TEST-NOT Arguments}{ adjoin&nset-difference&search\cr assoc&nset-exclusive-or&set-difference\cr count&nsublis&set-exclusive-or\cr delete&nsubst&sublis\cr delete-duplicates&nsubstitute&subsetp\cr find&nunion&subst\cr intersection&position&substitute\cr member&rassoc&tree-equal\cr mismatch&remove&union\cr nintersection&remove-duplicates&\cr }

\issue{EVAL-WHEN-NON-TOP-LEVEL:GENERALIZE-EVAL-NEW-KEYWORDS} \issue{EVAL-WHEN-OBSOLETE-KEYWORDS:X3J13-MAR-1993} The use of the situation names \misc{compile}, \misc{load}, and \misc{eval} in \specref{eval-when} is deprecated.

\endsubsection%{Deprecated Argument Conventions}

\beginsubsection{Deprecated Variables}

\issue{REQUIRE-PATHNAME-DEFAULTS-AGAIN:X3J13-DEC-91} The //[[CL:Glossary:variable]]// \varref{*modules*} is deprecated.

\endsubsection%{Deprecated Variables}

\beginsubsection{Deprecated Reader Syntax}

\issue{STRUCTURE-READ-PRINT-SYNTAX:KEYWORDS} The \f{\#S} //[[CL:Glossary:reader macro]]// forces keyword names into \thepackage{keyword}; \seesection\SharpsignS. This feature is deprecated; in the future, keyword names will be taken in the package they are read in, so //[[CL:Glossary:symbols]]// that are actually in \thepackage{keyword} should be used if that is what is desired.

\endsubsection%{Deprecated Reader Syntax}

