====== Function DRIBBLE ======

====Syntax====

**dribble** //''&optional'' pathname// → ////[[CL:Glossary:implementation-dependent]]////

====Arguments and Values====

//pathname// - a //[[CL:Glossary:pathname designator]]//.

====Description====

Either //[[CL:Glossary:binds]]// **[[CL:Variables:*standard-input*]]** and **[[CL:Variables:*standard-output*]]** or takes other appropriate action, so as to send a record of the input/output interaction to a file named by //pathname//. **[[CL:Functions:dribble]]** is intended to create a readable record of an interactive session.

If //pathname// is a //[[CL:Glossary:logical pathname]]//, it is translated into a physical pathname as if by calling **[[CL:Functions:translate-logical-pathname]]**.

''(dribble)'' terminates the recording of input and output and closes the dribble file.

If **[[CL:Functions:dribble]]** is //[[CL:Glossary:called]]// while a //[[CL:Glossary:stream]]// to a "dribble file" is still open from a previous //[[CL:Glossary:call]]// to **[[CL:Functions:dribble]]**, the effect is //[[CL:Glossary:implementation-defined]]//. For example, the already-//[[CL:Glossary:open]]// //[[CL:Glossary:stream]]// might be //[[CL:Glossary:closed]]//, or dribbling might occur both to the old //[[CL:Glossary:stream]]// and to a new one, or the old //[[CL:Glossary:stream]]// might stay open but not receive any further output, or the new request might be ignored, or some other action might be taken.

====Examples====

None.

====Affected By====

The //[[CL:Glossary:implementation]]//.

====Exceptional Situations====

If a failure occurs when performing some operation on the //[[CL:Glossary:file system]]// while creating the dribble file, an error of type **[[CL:Types:file-error]]** is signaled.

An error of type **[[CL:Types:file-error]]** might be signaled if //pathname// is a //[[CL:Glossary:designator]]// for a //[[CL:Glossary:wild]]// //[[CL:Glossary:pathname]]//.

====See Also====

{\secref\PathnamesAsFilenames}

====Notes====

**[[CL:Functions:dribble]]** can return before subsequent //[[CL:Glossary:forms]]// are executed. It also can enter a recursive interaction loop, returning only when ''(dribble)'' is done.

**[[CL:Functions:dribble]]** is intended primarily for interactive debugging; its effect cannot be relied upon when used in a program.

\issue{FILE-OPEN-ERROR:SIGNAL-FILE-ERROR} \issue{FILE-OPEN-ERROR:SIGNAL-FILE-ERROR} \issue{PATHNAME-HOST-PARSING:RECOGNIZE-LOGICAL-HOST-NAMES} \issue{DRIBBLE-TECHNIQUE} \issue{PATHNAME-LOGICAL:ADD} \issue{JUN90-TRIVIAL-ISSUES:25} \issue{FILE-OPEN-ERROR:SIGNAL-FILE-ERROR}
