====== Function DESCRIBE ======

====Syntax====

**describe {object** //\opt} stream// → //\novalues//

====Arguments and Values====

//object// - an //[[CL:Glossary:object]]//.

//stream// - an //[[CL:Glossary:output]]// //[[CL:Glossary:stream designator]]//. The default is //[[CL:Glossary:standard output]]//.

====Description====

**[[CL:Functions:describe]]** displays information about //object//

to //stream//.

For example, **[[CL:Functions:describe]]** of a //[[CL:Glossary:symbol]]// might show the //[[CL:Glossary:symbol]]//'s value, its definition, and each of its properties. **[[CL:Functions:describe]]** of a //[[CL:Glossary:float]]// might show the number's internal representation in a way that is useful for tracking down round-off errors. In all cases, however, the nature and format of the output of **[[CL:Functions:describe]]** is //[[CL:Glossary:implementation-dependent]]//.

**[[CL:Functions:describe]]** can describe something that it finds inside the //object//; in such cases, a notational device such as increased indentation or positioning in a table is typically used in order to visually distinguish such recursive descriptions from descriptions of the argument //object//.

The actual act of describing the object is implemented by **[[CL:Functions:describe-object]]**. **[[CL:Functions:describe]]** exists as an interface primarily to manage argument defaulting (including conversion of arguments \t\ and **[[CL:Constant Variables:nil]]** into //[[CL:Glossary:stream]]// //[[CL:Glossary:object|objects]]//) and to inhibit any return values from **[[CL:Functions:describe-object]]**.

**[[CL:Functions:describe]]** is not intended to be an interactive function. In a //[[CL:Glossary:conforming implementation]]//, **[[CL:Functions:describe]]** must not, by default, prompt for user input. User-defined methods for **[[CL:Functions:describe-object]]** are likewise restricted.

====Examples====

None.

====Side Effects====

Output to //[[CL:Glossary:standard output]]// or //[[CL:Glossary:terminal I/O]]//.

====Affected By====

**[[CL:Variables:*standard-output*]]** and **[[CL:Variables:*terminal-io*]]**, methods on **[[CL:Functions:describe-object]]** and **[[CL:Functions:print-object]]** for //[[CL:Glossary:object|objects]]// having user-defined //[[CL:Glossary:classes]]//.

====Exceptional Situations====

None.

====See Also====

**[[CL:Functions:inspect]]**, **[[CL:Functions:describe-object]]**

====Notes====

None.

\issue{DESCRIBE-UNDERSPECIFIED:DESCRIBE-OBJECT} \issue{DESCRIBE-UNDERSPECIFIED:DESCRIBE-OBJECT} \issue{DESCRIBE-UNDERSPECIFIED:DESCRIBE-OBJECT} \issue{DESCRIBE-INTERACTIVE:NO}
