====== Local Macro PPRINT-EXIT-IF-LIST-EXHAUSTED ======

====Syntax====

**pprint-exit-if-list-exhausted** //\noargs// → //**[[CL:Constant Variables:nil]]**//

====Arguments and Values====

None.

====Description====

Tests whether or not the //list// passed to the //[[CL:Glossary:lexically current logical block]]// has been exhausted; see section {\secref\DynamicControlofOutput}. If this //list// has been reduced to **[[CL:Constant Variables:nil]]**, **[[CL:Functions:pprint-exit-if-list-exhausted]]** terminates the execution of the //[[CL:Glossary:lexically current logical block]]// except for the printing of the suffix. Otherwise **[[CL:Functions:pprint-exit-if-list-exhausted]]** returns **[[CL:Constant Variables:nil]]**.

Whether or not **[[CL:Macros:pprint-exit-if-list-exhausted]]** is //[[CL:Glossary:fbound]]// in the //[[CL:Glossary:global environment]]// is //[[CL:Glossary:implementation-dependent]]//; however, the restrictions on redefinition and //[[CL:Glossary:shadowing]]// of **[[CL:Macros:pprint-exit-if-list-exhausted]]** are the same as for //[[CL:Glossary:symbols]]// in \thepackage{common-lisp} which are //[[CL:Glossary:fbound]]// in the //[[CL:Glossary:global environment]]//. The consequences of attempting to use **[[CL:Macros:pprint-exit-if-list-exhausted]]** outside of **[[CL:Macros:pprint-logical-block]]** are undefined.

====Examples====

None.

====Side Effects====

None.

====Affected By====

None.

====Exceptional Situations====

An error is signaled (at macro expansion time or at run time) if **[[CL:Functions:pprint-exit-if-list-exhausted]]** is used anywhere other than

lexically within a call on **[[CL:Macros:pprint-logical-block]]**.

Also, the consequences of executing **[[CL:Functions:pprint-if-list-exhausted]]** outside of the dynamic extent of the **[[CL:Macros:pprint-logical-block]]** which lexically contains it are undefined.

====See Also====

**[[CL:Macros:pprint-logical-block]]**, **[[CL:Functions:pprint-pop]]**.

====Notes====

None.

\issue{PRETTY-PRINT-INTERFACE} \issue{LEXICAL-CONSTRUCT-GLOBAL-DEFINITION:UNDEFINED}
