====== Function REDUCE ======

====Syntax====

\DefunWithValues reduce {function sequence ''&key'' key from-end start end initial-value} {result}

====Arguments and Values====

//function// - a //[[CL:Glossary:designator]]// for a //[[CL:Glossary:function]]// that might be called with either zero or two //[[CL:Glossary:arguments]]//.

//sequence// - a //[[CL:Glossary:proper sequence]]//.

//key// - a //[[CL:Glossary:designator]]// for a //[[CL:Glossary:function]]// of one argument, or **[[CL:Constant Variables:nil]]**.

//from-end// - a //[[CL:Glossary:generalized boolean]]//. The default is //[[CL:Glossary:false]]//.


//start//, //end// - //[[CL:Glossary:bounding index designators]]// of //sequence//. \Defaults{//start// and //end//}{''0'' and **[[CL:Constant Variables:nil]]**}

//initial-value// - an //[[CL:Glossary:object]]//.

//result// - an //[[CL:Glossary:object]]//.

====Description====

**[[CL:Functions:reduce]]** uses a binary operation, //function//, to combine the //[[CL:Glossary:element|elements]]// of //sequence// //[[CL:Glossary:bounded]]// by //start// and //end//.

The //function// must accept as //[[CL:Glossary:arguments]]// two //[[CL:Glossary:element|elements]]// of //sequence// or the results from combining those //[[CL:Glossary:element|elements]]//. The //function// must also be able to accept no arguments.

If //key// is supplied, it is used is used to extract the values to reduce. The //key// function is applied exactly once to each element of //sequence// in the order implied by the reduction order but not to the value of //initial-value//, if supplied.

The //key// function typically returns part of the //[[CL:Glossary:element]]// of //sequence//. If //key// is not supplied or is **[[CL:Constant Variables:nil]]**, the //sequence// //[[CL:Glossary:element]]// itself is used.

The reduction is left-associative, unless //from-end// is //[[CL:Glossary:true]]// in which case it is right-associative.

If //initial-value// is supplied, it is logically placed before the subsequence (or after it if //from-end// is //[[CL:Glossary:true]]//) and included in the reduction operation.

In the normal case, the result of **[[CL:Functions:reduce]]** is the combined result of //function//'s being applied to successive pairs of //[[CL:Glossary:element|elements]]// of //sequence//.

If the subsequence contains exactly one //[[CL:Glossary:element]]// and no //initial-value// is given, then that //[[CL:Glossary:element]]// is returned and //function// is not called. If the subsequence is empty and an //initial-value// is given, then the //initial-value// is returned and //function// is not called.

If the subsequence is empty and no //initial-value// is given, then the //function// is called with zero arguments, and **[[CL:Functions:reduce]]** returns whatever //function// does. This is the only case where the //function// is called with other than two arguments.

====Examples====

<blockquote> (reduce #'* '(1 2 3 4 5)) → 120 (reduce #'append '((1) (2)) :initial-value '(i n i t)) → (I N I T 1 2) (reduce #'append '((1) (2)) :from-end t :initial-value '(i n i t)) → (1 2 I N I T) (reduce #'- '(1 2 3 4)) ≡ (- (- (- 1 2) 3) 4) → -8 (reduce #'- '(1 2 3 4) :from-end t) ;Alternating sum. ≡ (- 1 (- 2 (- 3 4))) → -2 (reduce #'+ '()) → 0 (reduce #'+ '(3)) → 3 (reduce #'+ '(foo)) → FOO (reduce #'list '(1 2 3 4)) → (((1 2) 3) 4) (reduce #'list '(1 2 3 4) :from-end t) → (1 (2 (3 4))) (reduce #'list '(1 2 3 4) :initial-value 'foo) → ((((foo 1) 2) 3) 4) (reduce #'list '(1 2 3 4) :from-end t :initial-value 'foo) → (1 (2 (3 (4 foo)))) </blockquote>

====Side Effects====

None.

====Affected By====

None.

====Exceptional Situations====

Should be prepared to signal an error of type type-error if //sequence// is not a //[[CL:Glossary:proper sequence]]//.

====See Also====

{\secref\TraversalRules}

====Notes====

None.


\issue{MAPPING-DESTRUCTIVE-INTERACTION:EXPLICITLY-VAGUE} \issue{REDUCE-ARGUMENT-EXTRACTION} \issue{SUBSEQ-OUT-OF-BOUNDS} \issue{RANGE-OF-START-AND-END-PARAMETERS:INTEGER-AND-INTEGER-NIL} \issue{REDUCE-ARGUMENT-EXTRACTION}
