====== Function GET-UNIVERSAL-TIME, GET-DECODED-TIME ======

====Syntax====

**get-universal-time** //\noargs// → //universal-time//

\DefunWithValuesNewline get-decoded-time {\noargs} {second, minute, hour, date, month, year, day, daylight-p, zone}

====Arguments and Values====

//universal-time// - a //[[CL:Glossary:universal time]]//.

//second//, //minute//, //hour//, //date//, //month//, //year//, //day//, //daylight-p//, //zone// - a //[[CL:Glossary:decoded time]]//.

====Description====

**[[CL:Functions:get-universal-time]]** returns the current time, represented as a //[[CL:Glossary:universal time]]//.

**[[CL:Functions:get-decoded-time]]** returns the current time, represented as a //[[CL:Glossary:decoded time]]//.

====Examples====

<blockquote> ;; At noon on July 4, 1976 in Eastern Daylight Time. (get-decoded-time) → 0, 0, 12, 4, 7, 1976, 6, //[[CL:Glossary:true]]//, 5 ;; At exactly the same instant. (get-universal-time) → 2414332800 ;; Exactly five minutes later. (get-universal-time) → 2414333100 ;; The difference is 300 seconds (five minutes) (- * **) → 300 </blockquote>

====Side Effects====

None.

====Affected By====

The time of day (i.e. the passage of time), the system clock's ability to keep accurate time, and the accuracy of the system clock's initial setting.

====Exceptional Situations====

An error of type **[[CL:Types:error]]** might be signaled if the current time cannot be determined.

====See Also====

**[[CL:Functions:decode-universal-time]]**, **[[CL:Functions:encode-universal-time]]**, {\secref\Time}

====Notes====

<blockquote> (get-decoded-time) ≡ (decode-universal-time (get-universal-time)) </blockquote>

No //[[CL:Glossary:implementation]]// is required to have a way to verify that the time returned is correct. However, if an //[[CL:Glossary:implementation]]// provides a validity check (//e.g.// the failure to have properly initialized the system clock can be reliably detected) and that validity check fails, the //[[CL:Glossary:implementation]]// is strongly encouraged (but not required) to signal an error of type **[[CL:Types:error]]** (rather than, for example, returning a known-to-be-wrong value) that is //[[CL:Glossary:correctable]]// by allowing the user to interactively set the correct time.

