====== Function SUBSETP ======

====Syntax====

**subsetp {list-1 list-2** //\key} key test test-not// → //generalized-boolean//

====Arguments and Values====

//list-1// - a //[[CL:Glossary:proper list]]//.

//list-2// - a //[[CL:Glossary:proper list]]//.

//test// - a //[[CL:Glossary:designator]]// for a //[[CL:Glossary:function]]// of two //[[CL:Glossary:arguments]]// that returns a //[[CL:Glossary:generalized boolean]]//.

//test-not// - a //[[CL:Glossary:designator]]// for a //[[CL:Glossary:function]]// of two //[[CL:Glossary:arguments]]// that returns a //[[CL:Glossary:generalized boolean]]//.

//key// - a //[[CL:Glossary:designator]]// for a //[[CL:Glossary:function]]// of one argument, or **[[CL:Constant Variables:nil]]**.

//generalized-boolean// - a //[[CL:Glossary:generalized boolean]]//.

====Description====

**[[CL:Functions:subsetp]]** returns //[[CL:Glossary:true]]// if every element of //list-1// //matches// some element of //list-2//, and //[[CL:Glossary:false]]// otherwise.

Whether a list element is the same as another list element is determined by the functions specified by the keyword arguments. The first argument to the **'':test''** or **'':test-not''** function is

typically part of an element of //list-1// extracted by the **'':key''** function; the second argument is typically part of an element of //list-2// extracted by the **'':key''** function.

The argument to the **'':key''** function is an element of either //list-1// or //list-2//; the return value is part of the element of the supplied list element. If **'':key''** is not supplied or **[[CL:Constant Variables:nil]]**, the //list-1// or //list-2// element itself is supplied to the **'':test''** or **'':test-not''** function.

====Examples====

<blockquote> ([[CL:Macros:defparameter]] cosmos '(1 "a" (1 2))) → (1 "a" (1 2)) (subsetp '(1) cosmos) → //[[CL:Glossary:true]]// (subsetp '((1 2)) cosmos) → //[[CL:Glossary:false]]// (subsetp '((1 2)) cosmos :test 'equal) → //[[CL:Glossary:true]]// (subsetp '(1 "A") cosmos :test #'equalp) → //[[CL:Glossary:true]]// (subsetp '((1) (2)) '((1) (2))) → //[[CL:Glossary:false]]// (subsetp '((1) (2)) '((1) (2)) :key #'car) → //[[CL:Glossary:true]]// </blockquote>

====Side Effects====

None.

====Affected By====

None.

====Exceptional Situations====

Should be prepared to signal an error of type type-error if //list-1// and //list-2// are not //[[CL:Glossary:proper lists]]//.

====See Also====

{\secref\TraversalRules}

====Notes====

The **'':test-not''** parameter is deprecated.

\issue{MAPPING-DESTRUCTIVE-INTERACTION:EXPLICITLY-VAGUE} \issue{TEST-NOT-IF-NOT:FLUSH-ALL}
