====== Function INTERN ======

====Syntax====

**intern {string** //\opt} package// → //symbol, status//

====Arguments and Values====

//string// - a //[[CL:Glossary:string]]//.

//package// - a //[[CL:Glossary:package designator]]//.

The default is the //[[CL:Glossary:current package]]//.

//symbol// - a //[[CL:Glossary:symbol]]//.

//status// - one of **'':inherited''**, **'':external''**, **'':internal''**, or **[[CL:Constant Variables:nil]]**.

====Description====

**[[CL:Functions:intern]]** enters a //[[CL:Glossary:symbol]]// named //string// into //package//. If a //[[CL:Glossary:symbol]]// whose name is the same as //string// is already //[[CL:Glossary:accessible]]// in //package//, it is returned. If no such //[[CL:Glossary:symbol]]// is //[[CL:Glossary:accessible]]// in //package//, a new //[[CL:Glossary:symbol]]// with the given name is created and entered into //package// as an //[[CL:Glossary:internal symbol]]//, or as an //[[CL:Glossary:external symbol]]// if the //package// is \thepackage{keyword}; //package// becomes the //[[CL:Glossary:home package]]// of the created //[[CL:Glossary:symbol]]//.

The first value returned by **[[CL:Functions:intern]]**, //symbol//, is the //[[CL:Glossary:symbol]]// that was found or created.

The meaning of the //[[CL:Glossary:secondary value]]//, //status//, is as follows: \beginlist \itemitem{**'':internal''**}

The //[[CL:Glossary:symbol]]// was found and is

//[[CL:Glossary:present]]// in //package// as an //[[CL:Glossary:internal symbol]]//.

\itemitem{**'':external''**}

The //[[CL:Glossary:symbol]]// was found and is

//[[CL:Glossary:present]]// as an //[[CL:Glossary:external symbol]]//.

\itemitem{**'':inherited''**}

The //[[CL:Glossary:symbol]]// was found and is inherited via **[[CL:Functions:use-package]]** (which implies that the //[[CL:Glossary:symbol]]// is internal).

\itemitem{**[[CL:Constant Variables:nil]]**}

No pre-existing //[[CL:Glossary:symbol]]// was found, so one was created.

It is //[[CL:Glossary:implementation-dependent]]// whether the //[[CL:Glossary:string]]// that becomes the new //[[CL:Glossary:symbol]]//'s //[[CL:Glossary:name]]// is the given //string// or a copy of it. Once a //[[CL:Glossary:string]]// has been given as the //string// //[[CL:Glossary:argument]]// to //[[CL:Glossary:intern]]// in this situation where a new //[[CL:Glossary:symbol]]// is created, the consequences are undefined if a subsequent attempt is made to alter that //[[CL:Glossary:string]]//.

\endlist

====Examples====

<blockquote> (in-package "COMMON-LISP-USER") → #<PACKAGE "COMMON-LISP-USER"> (intern "Never-Before") → |Never-Before|, NIL (intern "Never-Before") → |Never-Before|, :INTERNAL (intern "NEVER-BEFORE" "KEYWORD") → :NEVER-BEFORE, NIL (intern "NEVER-BEFORE" "KEYWORD") → :NEVER-BEFORE, :EXTERNAL </blockquote>

====Side Effects====

None.

====Affected By====

None.

====Exceptional Situations====

None.

====See Also====

**[[CL:Functions:find-symbol]]**, **[[CL:Functions:read]]**, **[[CL:Types:symbol]]**, **[[CL:Functions:unintern]]**, {\secref\SymbolTokens}

====Notes====

**[[CL:Functions:intern]]** does not need to do any name conflict checking because it never creates a new //[[CL:Glossary:symbol]]// if there is already an //[[CL:Glossary:accessible]]// //[[CL:Glossary:symbol]]// with the name given.

\issue{PACKAGE-FUNCTION-CONSISTENCY:MORE-PERMISSIVE} \issue{PACKAGE-FUNCTION-CONSISTENCY:MORE-PERMISSIVE}
