====== Function APROPOS, APROPOS-LIST ======

====Syntax====

**apropos {string** //\opt} package// → //\novalues//

**apropos-list {string** //\opt} package// → //symbols//

====Arguments and Values====

//string// - a //[[CL:Glossary:\symbolnamedesignator]]//.

//package// - a //[[CL:Glossary:package designator]]// or **[[CL:Constant Variables:nil]]**. The default is **[[CL:Constant Variables:nil]]**.

//symbols// - a //[[CL:Glossary:list]]// of //[[CL:Glossary:symbols]]//.

====Description====

These functions search for //[[CL:Glossary:interned]]// //[[CL:Glossary:symbols]]// whose //[[CL:Glossary:names]]// contain the substring //string//.

For **[[CL:Functions:apropos]]**, as each such //[[CL:Glossary:symbol]]// is found, its name is printed on //[[CL:Glossary:standard output]]//. In addition, if such a //[[CL:Glossary:symbol]]// is defined as a //[[CL:Glossary:function]]// or //[[CL:Glossary:dynamic variable]]//, information about those definitions might also be printed.

For **[[CL:Functions:apropos-list]]**, no output occurs as the search proceeds; instead a list of the matching //[[CL:Glossary:symbols]]// is returned when the search is complete.

If //package// is //[[CL:Glossary:non-nil]]//, only the //[[CL:Glossary:symbols]]// //[[CL:Glossary:accessible]]// in that //package// are searched; otherwise all //[[CL:Glossary:symbols]]// //[[CL:Glossary:accessible]]// in any //[[CL:Glossary:package]]// are searched.

Because a //[[CL:Glossary:symbol]]// might be available by way of more than one inheritance path, **[[CL:Functions:apropos]]** might print information about the //[[CL:Glossary:same]]// //[[CL:Glossary:symbol]]// more than once,

or **[[CL:Functions:apropos-list]]** might return a //[[CL:Glossary:list]]// containing duplicate //[[CL:Glossary:symbols]]//.

Whether or not the search is case-sensitive is //[[CL:Glossary:implementation-defined]]//.

====Examples====

None.

====Affected By====

The set of //[[CL:Glossary:symbols]]// which are currently //[[CL:Glossary:interned]]// in any //[[CL:Glossary:packages]]// being searched.

**[[CL:Functions:apropos]]** is also affected by **[[CL:Variables:*standard-output*]]**.

====Exceptional Situations====

None.

====See Also====

None.

====Notes====

None.

