====== Function ENSURE-DIRECTORIES-EXIST ======

====Syntax====

**ensure-directories-exist {pathspec** //\key} verbose// → //pathspec, created//

====Arguments and Values====

//pathspec// - a //[[CL:Glossary:pathname designator]]//.

//verbose// - a //[[CL:Glossary:generalized boolean]]//.

//created// - a //[[CL:Glossary:generalized boolean]]//.

====Description====

Tests whether the directories containing the specified //[[CL:Glossary:file]]// actually exist, and attempts to create them if they do not.

If the containing directories do not exist and if //verbose// is //[[CL:Glossary:true]]//, then the //[[CL:Glossary:implementation]]// is permitted (but not required) to perform output to //[[CL:Glossary:standard output]]// saying what directories were created. If the containing directories exist, or if //verbose// is //[[CL:Glossary:false]]//, this function performs no output.

The //[[CL:Glossary:primary value]]// is the given //[[CL:Glossary:pathspec]]// so that this operation can be straightforwardly composed with other file manipulation expressions. The //[[CL:Glossary:secondary value]]//, //created//, is //[[CL:Glossary:true]]// if any directories were created.

====Examples====

None.

====Affected By====

The host computer's file system.

====Exceptional Situations====

An error of type **[[CL:Types:file-error]]** is signaled if the host, device, or directory part of //pathspec// is //[[CL:Glossary:wild]]//.

If the directory creation attempt is not successful, an error of type **[[CL:Types:file-error]]** is signaled; if this occurs, it might be the case that none, some, or all of the requested creations have actually occurred within the //[[CL:Glossary:file system]]//.

====See Also====

**[[CL:Functions:probe-file]]**, **[[CL:Functions:open]]**,

{\secref\PathnamesAsFilenames}

====Notes====

None.

\issue{FILE-OPEN-ERROR:SIGNAL-FILE-ERROR} \issue{FILE-OPEN-ERROR:SIGNAL-FILE-ERROR} \issue{FILE-OPEN-ERROR:SIGNAL-FILE-ERROR}
