====== Function INVALID-METHOD-ERROR ======

====Syntax====

\DefunWithValues invalid-method-error {method format-control ''&rest'' args} {//[[CL:Glossary:implementation-dependent]]//}

====Arguments and Values====

//method// - a //[[CL:Glossary:method]]//.


//format-control// - a //[[CL:Glossary:format control]]//.


//args// - //[[CL:Glossary:format arguments]]// for the //format-control//.

====Description====

The //[[CL:Glossary:function]]// **[[CL:Functions:invalid-method-error]]** is used to signal an error of type **[[CL:Types:error]]** when there is an applicable //[[CL:Glossary:method]]// whose //[[CL:Glossary:qualifiers]]// are not valid for the method combination type. The error message is constructed by using the //format-control// suitable for **[[CL:Functions:format]]** and any //args// to it. Because an implementation may need to add additional contextual information to the error message, **[[CL:Functions:invalid-method-error]]** should be called only within the dynamic extent of a method combination function.

The //[[CL:Glossary:function]]// **[[CL:Functions:invalid-method-error]]** is called automatically when a //[[CL:Glossary:method]]// fails to satisfy every //[[CL:Glossary:qualifier]]// pattern and predicate in a **[[CL:Macros:define-method-combination]]** //[[CL:Glossary:form]]//. A method combination function that imposes additional restrictions should call **[[CL:Functions:invalid-method-error]]** explicitly if it encounters a //[[CL:Glossary:method]]// it cannot accept.


Whether **[[CL:Functions:invalid-method-error]]** returns to its caller or exits via \specref{throw} is //[[CL:Glossary:implementation-dependent]]//.

====Examples====

None.

====Side Effects====

The debugger might be entered.

====Affected By====

**[[CL:Variables:*break-on-signals*]]**

====Exceptional Situations====

None.

====See Also====

**[[CL:Macros:define-method-combination]]**


====Notes====

None.


\issue{FORMAT-STRING-ARGUMENTS:SPECIFY}
