====== Function WARN ======

====Syntax====

**warn {datum** //\rest} arguments// → //**[[CL:Constant Variables:nil]]**//

====Arguments and Values====

//datum//, //arguments// - //[[CL:Glossary:designators]]// for a //[[CL:Glossary:condition]]// of default type **[[CL:Types:simple-warning]]**.

====Description====

//[[CL:Glossary:Signals]]// a //[[CL:Glossary:condition]]// of type **[[CL:Types:warning]]**. If the //[[CL:Glossary:condition]]// is not //[[CL:Glossary:handled]]//, reports the //[[CL:Glossary:condition]]// to //[[CL:Glossary:error output]]//.

The precise mechanism for warning is as follows:

\beginlist

\itemitem{{\bf The warning condition is signaled}}

While the **[[CL:Types:warning]]** //[[CL:Glossary:condition]]// is being signaled, \therestart{muffle-warning} is established for use by a //[[CL:Glossary:handler]]//. If invoked, this //[[CL:Glossary:restart]]// bypasses further action by **[[CL:Functions:warn]]**, which in turn causes **[[CL:Functions:warn]]** to immediately return **[[CL:Constant Variables:nil]]**.

\itemitem{{\bf If no handler for the warning condition is found}}

If no handlers for the warning condition are found, or if all such handlers decline, then the //[[CL:Glossary:condition]]// is reported to //[[CL:Glossary:error output]]// by **[[CL:Functions:warn]]** in an //[[CL:Glossary:implementation-dependent]]// format.

\itemitem{{\bf **[[CL:Constant Variables:nil]]** is returned}}

The value returned by **[[CL:Functions:warn]]** if it returns is **[[CL:Constant Variables:nil]]**. \endlist

====Examples====

<blockquote> (defun foo (x) (let ((result (* x 2))) (if (not (typep result 'fixnum)) (warn "You're using very big numbers.")) result)) → FOO

(foo 3) → 6

(foo most-positive-fixnum)
▷ Warning: You're using very big numbers. → 4294967294

([[CL:Macros:defparameter]] *break-on-signals* t) → T

(foo most-positive-fixnum)
▷ Break: Caveat emptor.
▷ To continue, type :CONTINUE followed by an option number.
▷ 1: Return from Break.
▷ 2: Abort to Lisp Toplevel.
▷ Debug> :continue 1
▷ Warning: You're using very big numbers. → 4294967294 </blockquote>

====Side Effects====

A warning is issued. The debugger might be entered.

====Affected By====

Existing handler bindings.

**[[CL:Variables:*break-on-signals*]]**, **[[CL:Variables:*error-output*]]**.

====Exceptional Situations====

If //datum// is a //[[CL:Glossary:condition]]// and if the //[[CL:Glossary:condition]]// is not of type **[[CL:Types:warning]]**, or //arguments// is //[[CL:Glossary:non-nil]]//, an error of type **[[CL:Types:type-error]]** is signaled.

If //datum// is a condition type, the result of **[[CL:Functions:(apply #'make-condition datum arguments)]]** must be of type **[[CL:Types:warning]]** or an error of type **[[CL:Types:type-error]]** is signaled.

====See Also====

**[[CL:Variables:*break-on-signals*]]**, **[[CL:Functions:muffle-warning]]**, **[[CL:Functions:signal]]**

====Notes====

None.

\issue{BREAK-ON-WARNINGS-OBSOLETE:REMOVE}
