====== Function LOAD-LOGICAL-PATHNAME-TRANSLATIONS ======

====Syntax====

**load-logical-pathname-translations** //host// → //just-loaded//

====Arguments and Values====

//host// - a //[[CL:Glossary:string]]//.

//[[CL:Glossary:just-loaded]]//---a //[[CL:Glossary:generalized boolean]]//.

====Description====

Searches for and loads the definition of a //[[CL:Glossary:logical host]]// named //host//, if it is not already defined. The specific nature of the search is //[[CL:Glossary:implementation-defined]]//.

If the //host// is already defined, no attempt to find or load a definition is attempted, and //[[CL:Glossary:false]]// is returned. If the //host// is not already defined, but a definition is successfully found and loaded, //[[CL:Glossary:true]]// is returned. Otherwise, an error is signaled.

====Examples====

<blockquote> (translate-logical-pathname "hacks:weather;barometer.lisp.newest")
▷ Error: The logical host HACKS is not defined. (load-logical-pathname-translations "HACKS")
▷ ;; Loading SYS:SITE;HACKS.TRANSLATIONS
▷ ;; Loading done. → //[[CL:Glossary:true]]// (translate-logical-pathname "hacks:weather;barometer.lisp.newest") → #P"HELIUM:[SHARED.HACKS.WEATHER]BAROMETER.LSP;0" (load-logical-pathname-translations "HACKS") → //[[CL:Glossary:false]]// </blockquote>

====Affected By====

None.

====Exceptional Situations====

If no definition is found, an error of type **[[CL:Types:error]]** is signaled.

====See Also====

**[[CL:Types:logical-pathname]]**

====Notes====

//[[CL:Glossary:Logical pathname]]// definitions will be created not just by //[[CL:Glossary:implementors]]// but also by //[[CL:Glossary:programmers]]//. As such, it is important that the search strategy be documented. For example, an //[[CL:Glossary:implementation]]// might define that the definition of a //host// is to be found in a file called "//host//.translations" in some specifically named directory.


\issue{PATHNAME-LOGICAL:ADD}
