====== Function MAX, MIN ======

====Syntax====

**max** //''&rest'' reals''+''// → //max-real// **min** //''&rest'' reals''+''// → //min-real//

====Arguments and Values====

//real// - a //[[CL:Glossary:real]]//.

//max-real//, //min-real// - a //[[CL:Glossary:real]]//.

====Description====

**[[CL:Functions:max]]** returns the //real// that is greatest (closest to positive infinity). **[[CL:Functions:min]]** returns the //real// that is least (closest to negative infinity).

For **[[CL:Functions:max]]**, the implementation has the choice of returning the largest argument as is or applying the rules of floating-point //[[CL:Glossary:contagion]]//, taking all the arguments into consideration for //[[CL:Glossary:contagion]]// purposes. Also, if one or more of the arguments are **[[CL:Functions:=]]**, then any one of them may be chosen as the value to return. For example, if the //reals// are a mixture of //[[CL:Glossary:rationals]]// and //[[CL:Glossary:floats]]//, and the largest argument is a //[[CL:Glossary:rational]]//, then the implementation is free to produce either that //[[CL:Glossary:rational]]// or its //[[CL:Glossary:float]]// approximation; if the largest argument is a //[[CL:Glossary:float]]// of a smaller format than the largest format of any //[[CL:Glossary:float]]// argument, then the implementation is free to return the argument in its given format or expanded to the larger format. Similar remarks apply to **[[CL:Functions:min]]** (replacing "largest argument" by "smallest argument").

====Examples====

<blockquote> (max 3) → 3 (min 3) → 3 (max 6 12) → 12 (min 6 12) → 6 (max -6 -12) → -6 (min -6 -12) → -12 (max 1 3 2 -7) → 3 (min 1 3 2 -7) → -7 (max -2 3 0 7) → 7 (min -2 3 0 7) → -2 (max 5.0 2) → 5.0 (min 5.0 2) → 2 //or// → 2.0 (max 3.0 7 1) → 7 //or// → 7.0 (min 3.0 7 1) → 1 //or// → 1.0 (max 1.0s0 7.0d0) → 7.0d0 (min 1.0s0 7.0d0) → 1.0s0 //or// → 1.0d0 (max 3 1 1.0s0 1.0d0) → 3 //or// → 3.0d0 (min 3 1 1.0s0 1.0d0) → 1 //or// → 1.0s0 //or// → 1.0d0 </blockquote>

====Side Effects====

None.

====Affected By====

None.

====Exceptional Situations====

\Shouldcheckanytype{number}{a //[[CL:Glossary:real]]//}

====See Also====

None.

====Notes====

None.



