====== Function UNINTERN ======

====Syntax====

**unintern {symbol** //\opt} package// → //generalized-boolean//

====Arguments and Values====

//symbol// - a //[[CL:Glossary:symbol]]//.

//package// - a //[[CL:Glossary:package designator]]//.

The default is the //[[CL:Glossary:current package]]//.

//generalized-boolean// - a //[[CL:Glossary:generalized boolean]]//.

====Description==== **[[CL:Functions:unintern]]** removes //symbol// from //package//.

If //symbol// is //[[CL:Glossary:present]]// in //package//, it is removed from //package// and also from //package//'s //[[CL:Glossary:shadowing symbols list]]// if it is present there. If //package// is the //[[CL:Glossary:home package]]// for //symbol//, //symbol// is made to have no //[[CL:Glossary:home package]]//. //Symbol// may continue to be //[[CL:Glossary:accessible]]// in //package// by inheritance.

Use of **[[CL:Functions:unintern]]** can result in a //[[CL:Glossary:symbol]]// that has no recorded //[[CL:Glossary:home package]]//, but that in fact is //[[CL:Glossary:accessible]]// in some //[[CL:Glossary:package]]//. \clisp\ does not check for this pathological case, and such //[[CL:Glossary:symbols]]// are always printed preceded by ''#:''.

**[[CL:Functions:unintern]]** returns //[[CL:Glossary:true]]// if it removes //symbol//, and **[[CL:Constant Variables:nil]]** otherwise.

====Examples====

<blockquote> (in-package "COMMON-LISP-USER") → #<PACKAGE "COMMON-LISP-USER"> ([[CL:Macros:defparameter]] temps-unpack (intern "UNPACK" (make-package 'temp))) → TEMP::UNPACK (unintern temps-unpack 'temp) → T (find-symbol "UNPACK" 'temp) → NIL, NIL temps-unpack → #:UNPACK </blockquote>

====Side Effects====

**[[CL:Functions:unintern]]** changes the state of the package system in such a way that the consistency rules do not hold across the change.

====Affected By==== Current state of the package system.

====Exceptional Situations====

Giving a shadowing symbol to **[[CL:Functions:unintern]]** can uncover a name conflict that had previously been resolved by the shadowing. If package A uses packages B and C, A contains a shadowing symbol ''x'', and B and C each contain external symbols named ''x'', then removing the shadowing symbol ''x'' from A will reveal a name conflict between ''b:x'' and ''c:x'' if those two //[[CL:Glossary:symbols]]// are distinct. In this case **[[CL:Functions:unintern]]** will signal an error.

====See Also====

{\secref\PackageConcepts}

====Notes====

None.

\issue{PACKAGE-FUNCTION-CONSISTENCY:MORE-PERMISSIVE}
