====== Accessor SUBSEQ ======

====Syntax====

**subseq {sequence start** //\opt} end// → //subsequence// (**setf** (**subseq {sequence start** //\opt} end//) //new-subsequence//)

====Arguments and Values====

//sequence// - a //[[CL:Glossary:proper sequence]]//.

//start//, //end// - //[[CL:Glossary:bounding index designators]]// of //sequence//. \DefaultFor{//end//}{**[[CL:Constant Variables:nil]]**}

//subsequence// - a //[[CL:Glossary:proper sequence]]//.

//new-subsequence// - a //[[CL:Glossary:proper sequence]]//.

====Description====

**[[CL:Functions:subseq]]** creates a //[[CL:Glossary:sequence]]// that is a copy of the subsequence of //sequence// //bounded// by //start// and //end//.

//Start// specifies an offset into the original //sequence// and marks the beginning position of the subsequence. //end// marks the position following the last element of the subsequence.

**[[CL:Functions:subseq]]** always allocates a new //[[CL:Glossary:sequence]]// for a result; it never shares storage with an old //[[CL:Glossary:sequence]]//. The result subsequence is always of the same //[[CL:Glossary:type]]// as //sequence//.

If //sequence// is a //[[CL:Glossary:vector]]//, the result is a //[[CL:Glossary:fresh]]// //[[CL:Glossary:simple array]]// of //[[CL:Glossary:rank]]// one that has the same //[[CL:Glossary:actual array element type]]// as //sequence//. If //sequence// is a //[[CL:Glossary:list]]//, the result is a //[[CL:Glossary:fresh]]// //[[CL:Glossary:list]]//.

**[[CL:Macros:setf]]** may be used with **[[CL:Functions:subseq]]** to destructively replace //[[CL:Glossary:element|elements]]// of a subsequence with //[[CL:Glossary:element|elements]]// taken from a //[[CL:Glossary:sequence]]// of new values. If the subsequence and the new sequence are not of equal length, the shorter length determines the number of elements that are replaced. The remaining //[[CL:Glossary:element|elements]]// at the end of the longer sequence are not modified in the operation.

====Examples====

<blockquote> ([[CL:Macros:defparameter]] str "012345") → "012345" (subseq str 2) → "2345" (subseq str 3 5) → "34" ([[CL:Macros:setf]] (subseq str 4) "abc") → "abc" str → "0123ab" ([[CL:Macros:setf]] (subseq str 0 2) "A") → "A" str → "A123ab" </blockquote>

====Side Effects====

None.

====Affected By====

None.

====Exceptional Situations====

Should be prepared to signal an error of type type-error if //sequence// is not a //[[CL:Glossary:proper sequence]]//. Should be prepared to signal an error of type type-error if //new-subsequence// is not a //[[CL:Glossary:proper sequence]]//.

====See Also====

**[[CL:Functions:replace]]**

====Notes====

None.

\issue{SUBSEQ-OUT-OF-BOUNDS} \issue{RANGE-OF-START-AND-END-PARAMETERS:INTEGER-AND-INTEGER-NIL}
