====== Function WILD-PATHNAME-P ======

====Syntax====

**wild-pathname-p {pathname** //\opt} field-key// → //generalized-boolean//

====Arguments and Values====

//pathname// - a //[[CL:Glossary:pathname designator]]//.

//Field-key// - one of **'':host''**, **'':device''** **'':directory''**, **'':name''**, **'':type''**, **'':version''**, or **[[CL:Constant Variables:nil]]**.

//generalized-boolean// - a //[[CL:Glossary:generalized boolean]]//.

====Description====

**[[CL:Functions:wild-pathname-p]]** tests //pathname// for the presence of wildcard components.

If //pathname// is a //[[CL:Glossary:pathname]]// (as returned by **[[CL:Functions:pathname]]**) it represents the name used to open the file. This may be, but is not required to be, the actual name of the file.

If //field-key// is not supplied or **[[CL:Constant Variables:nil]]**, **[[CL:Functions:wild-pathname-p]]** returns true if //pathname// has any wildcard components, **[[CL:Constant Variables:nil]]** if //pathname// has none. If //field-key// is //[[CL:Glossary:non-nil]]//, **[[CL:Functions:wild-pathname-p]]** returns true if the indicated component of //pathname// is a wildcard, **[[CL:Constant Variables:nil]]** if the component is not a wildcard.

====Examples==== <blockquote> ;;;The following examples are not portable. They are written to run ;;;with particular file systems and particular wildcard conventions. ;;;Other implementations will behave differently. These examples are ;;;intended to be illustrative, not to be prescriptive.

(wild-pathname-p (make-pathname :name :wild)) → //[[CL:Glossary:true]]// (wild-pathname-p (make-pathname :name :wild) :name) → //[[CL:Glossary:true]]// (wild-pathname-p (make-pathname :name :wild) :type) → //[[CL:Glossary:false]]// (wild-pathname-p (pathname "s:>foo>**>")) → //[[CL:Glossary:true]]// ;Lispm (wild-pathname-p (pathname :name "F*O")) → //[[CL:Glossary:true]]// ;Most places </blockquote>

====Affected By====

None.

====Exceptional Situations====

If //pathname// is not a //[[CL:Glossary:pathname]]//, a //[[CL:Glossary:string]]//, or a //[[CL:Glossary:stream associated with a file]]// an error of type **[[CL:Types:type-error]]** is signaled.

====See Also====

**[[CL:Types:pathname]]**, **[[CL:Types:logical-pathname]]**,{\secref\FileSystemConcepts},

{\secref\PathnamesAsFilenames}

====Notes====

Not all implementations support wildcards in all fields. see section {\secref\WildComponents}\ and \secref\WildcardRestrictions.


\issue{PATHNAME-WILD:NEW-FUNCTIONS} \issue{PATHNAME-STREAM} \issue{PATHNAME-LOGICAL:ADD} \issue{FILE-OPEN-ERROR:SIGNAL-FILE-ERROR}
