====== Function ED ======

====Syntax====

**ed** //''&optional'' x// → ////[[CL:Glossary:implementation-dependent]]////

====Arguments and Values====

//x// - **[[CL:Constant Variables:nil]]**, a //[[CL:Glossary:pathname]]//, a //[[CL:Glossary:string]]//, or a //[[CL:Glossary:function name]]//.

The default is **[[CL:Constant Variables:nil]]**.

====Description====

**[[CL:Functions:ed]]** invokes the editor if the //[[CL:Glossary:implementation]]// provides a resident editor.

If //x// is **[[CL:Constant Variables:nil]]**, the editor is entered. If the editor had been previously entered, its prior state is resumed, if possible.

If //x// is a //[[CL:Glossary:pathname]]// or //[[CL:Glossary:string]]//, it is taken as the //[[CL:Glossary:pathname designator]]// for a //[[CL:Glossary:file]]// to be edited.

If //x// is a //[[CL:Glossary:function name]]//, the text of its definition is edited. The means by which the function text is obtained is //[[CL:Glossary:implementation-defined]]//.

====Examples====

None.

====Affected By====

None.

====Exceptional Situations====

The consequences are undefined if the //[[CL:Glossary:implementation]]// does not provide a resident editor.

Might signal **[[CL:Types:type-error]]** if its argument is supplied but is not a //[[CL:Glossary:symbol]]//, a //[[CL:Glossary:pathname]]//, or **[[CL:Constant Variables:nil]]**.

If a failure occurs when performing some operation on the //[[CL:Glossary:file system]]// while attempting to edit a //[[CL:Glossary:file]]//, an error of type **[[CL:Types:file-error]]** is signaled.

An error of type **[[CL:Types:file-error]]** might be signaled if //x// is a //[[CL:Glossary:designator]]// for a //[[CL:Glossary:wild]]// //[[CL:Glossary:pathname]]//.

//[[CL:Glossary:Implementation-dependent]]// additional conditions might be signaled as well.

====See Also====

**[[CL:Types:pathname]]**, **[[CL:Types:logical-pathname]]**, **[[CL:Functions:compile-file]]**, **[[CL:Functions:load]]**,

{\secref\PathnamesAsFilenames}

====Notes====

None.



\issue{FUNCTION-NAME:LARGE} \issue{FILE-OPEN-ERROR:SIGNAL-FILE-ERROR} \issue{FILE-OPEN-ERROR:SIGNAL-FILE-ERROR} \issue{PATHNAME-LOGICAL:ADD} \issue{FILE-OPEN-ERROR:SIGNAL-FILE-ERROR} \issue{PATHNAME-HOST-PARSING:RECOGNIZE-LOGICAL-HOST-NAMES}
