====== Function PPRINT-INDENT ======

====Syntax====

**pprint-indent {relative-to n** //\opt} stream// → //**[[CL:Constant Variables:nil]]**//

====Arguments and Values====

//relative-to// - either **'':block''** or **'':current''**.

//n// - a //[[CL:Glossary:real]]//.

//stream// - an //[[CL:Glossary:output]]// //[[CL:Glossary:stream designator]]//. The default is //[[CL:Glossary:standard output]]//.

====Description====

**[[CL:Functions:pprint-indent]]** specifies the indentation to use in a logical block on //[[CL:Glossary:stream]]//.

If //stream// is a //[[CL:Glossary:pretty printing stream]]// and \thevalueof{*print-pretty*} is //[[CL:Glossary:true]]//, **[[CL:Functions:pprint-indent]]** sets the indentation in the innermost dynamically enclosing logical block; otherwise, **[[CL:Functions:pprint-indent]]** has no effect.

//N// specifies the indentation in //[[CL:Glossary:ems]]//. If //relative-to// is **'':block''**, the indentation is set to the horizontal position of the first character in the //[[CL:Glossary:dynamically current logical block]]// plus //n// //[[CL:Glossary:ems]]//. If //relative-to// is **'':current''**, the indentation is set to the current output position plus //n// //[[CL:Glossary:ems]]//. (For robustness in the face of variable-width fonts, it is advisable to use **'':current''** with an //n// of zero whenever possible.)

//N// can be negative; however, the total indentation cannot be moved left of the beginning of the line or left of the end of the rightmost per-line prefix---an attempt to move beyond one of these limits is treated

the same as an attempt to move to that limit. Changes in indentation caused by //pprint-indent// do not take effect until after the next line break. In addition, in miser mode all calls to **[[CL:Functions:pprint-indent]]** are ignored, forcing the lines corresponding to the logical block to line up under the first character in the block.

====Examples====

None.

====Side Effects====

None.

====Affected By====

None.

====Exceptional Situations====

An error is signaled if //relative-to// is any //[[CL:Glossary:object]]// other than **'':block''** or **'':current''**.

====See Also====

{\secref\TildeI}

====Notes====

None.

\issue{PRETTY-PRINT-INTERFACE} \issue{GENERALIZE-PRETTY-PRINTER:UNIFY}
